---
title: "Datacamp_Correlation and Regression__Model Fit"
author: "dizhen"
date: "2019Äê4ÔÂ18ÈÕ"
output: html_document
---


# Standard error of residuals
One way to assess strength of fit is to consider how far off the model is for a typical case. That is, for some observations, the fitted value will be very close to the actual value, while for others it will not. The magnitude of a typical residual can give us a sense of generally how close our estimates are.

However, recall that some of the residuals are positive, while others are negative. In fact, it is guaranteed by the least squares fitting procedure that the mean of the residuals is zero. Thus, it makes more sense to compute the square root of the mean squared residual, or root mean squared error (RMSE). R calls this quantity the residual standard error.

To make this estimate unbiased, you have to divide the sum of the squared residuals by the degrees of freedom in the model. Thus,

You can recover the residuals from mod with residuals(), and the degrees of freedom with df.residual().

Instructions

View a summary() of mod.
Compute the mean of the residuals() and verify that it is approximately zero.
Use residuals() and df.residual() to compute the root mean squared error (RMSE), a.k.a. residual standard error.

```{r}
# View summary of model
summary(mod)

# Compute the mean of the residuals
mean(residuals(mod))

# Compute RMSE
sqrt(sum(residuals(mod)^2) / df.residual(mod))
```

# Assessing simple linear model fit
Recall that the coefficient of determination (R2), can be computed as

where e is the vector of residuals and y is the response variable. This gives us the interpretation of R2 as the percentage of the variability in the response that is explained by the model, since the residuals are the part of that variability that remains unexplained by the model.

Instructions

The bdims_tidy data frame is the result of augment()-ing the bdims data frame with the mod for wgt as a function of hgt.

Use the summary() function to view the full results of mod.
Use the bdims_tidy data frame to compute the R2 of mod manually using the formula above, by computing the ratio of the variance of the residuals to the variance of the response variable.

```{r}
# View model summary
summary(mod)

# Compute R-squared
bdims_tidy %>%
  summarize(var_y = var(wgt), var_e = var(.resid)) %>%
  mutate(R_squared = 1 - var_e/var_y)
```
# Linear vs. average
The R2 gives us a numerical measurement of the strength of fit relative to a null model based on the average of the response variable:


This model has an R2 of zero because SSE=SST. That is, since the fitted values (y^null) are all equal to the average (y¡¥), the residual for each observation is the distance between that observation and the mean of the response. Since we can always fit the null model, it serves as a baseline against which all other models will be compared.

In the graphic, we visualize the residuals for the null model (mod_null at left) vs. the simple linear regression model (mod_hgt at right) with height as a single explanatory variable. Try to convince yourself that, if you squared the lengths of the grey arrows on the left and summed them up, you would get a larger value than if you performed the same operation on the grey arrows on the right.

It may be useful to preview these augment()-ed data frames with glimpse():

```{r}
glimpse(mod_null)
glimpse(mod_hgt)
```

Instructions

Compute the sum of the squared residuals (SSE) for the null model mod_null.
Compute the sum of the squared residuals (SSE) for the regression model mod_hgt.

```{r}
# Compute SSE for null model
mod_null %>%
  summarize(SSE = var(.resid))

# Compute SSE for regression model
mod_hgt %>%
  summarize(SSE = var(.resid))
```
# Leverage
The leverage of an observation in a regression model is defined entirely in terms of the distance of that observation from the mean of the explanatory variable. That is, observations close to the mean of the explanatory variable have low leverage, while observations far from the mean of the explanatory variable have high leverage. Points of high leverage may or may not be influential.

The augment() function from the broom package will add the leverage scores (.hat) to a model data frame.

Instructions

Use augment() to list the top 6 observations by their leverage scores, in descending order.

```{r}
# Rank points of high leverage
mod %>%
  augment() %>%
  arrange(desc(.hat)) %>%
  head()
```

# Influence
As noted previously, observations of high leverage may or may not be influential. The influence of an observation depends not only on its leverage, but also on the magnitude of its residual. Recall that while leverage only takes into account the explanatory variable (x), the residual depends on the response variable (y) and the fitted value (y^).

Influential points are likely to have high leverage and deviate from the general relationship between the two variables. We measure influence using Cook's distance, which incorporates both the leverage and residual of each observation.

Instructions

Use augment() to list the top 6 observations by their Cook's distance (.cooksd), in descending order.

```{r}
# Rank influential points
mod %>%
  augment() %>%
  arrange(desc(.cooksd)) %>%
  head()
```

# Removing outliers
Observations can be outliers for a number of different reasons. Statisticians must always be careful¡ªand more importantly, transparent¡ªwhen dealing with outliers. Sometimes, a better model fit can be achieved by simply removing outliers and re-fitting the model. However, one must have strong justification for doing this. A desire to have a higher R2 is not a good enough reason!

In the mlbBat10 data, the outlier with an OBP of 0.550 is Bobby Scales, an infielder who had four hits in 13 at-bats for the Chicago Cubs. Scales also walked seven times, resulting in his unusually high OBP. The justification for removing Scales here is weak. While his performance was unusual, there is nothing to suggest that it is not a valid data point, nor is there a good reason to think that somehow we will learn more about Major League Baseball players by excluding him.

Nevertheless, we can demonstrate how removing him will affect our model.

Instructions

Use filter() to create a subset of mlbBat10 called nontrivial_players consisting of only those players with at least 10 at-bats and OBP of below 0.500.
Fit the linear model for SLG as a function of OBP for the nontrivial_players. Save the result as mod_cleaner.
View the summary() of the new model and compare the slope and R2 to those of mod, the original model fit to the data on all players.
Visualize the new model with ggplot() and the appropriate geom_*() functions.

```{r}
# Create nontrivial_players
nontrivial_players <- mlbBat10 %>%
  filter(AB >= 10, OBP < 0.5)

# Fit model to new data
mod_cleaner <- lm(SLG ~ OBP, data = nontrivial_players)

# View model summary
summary(mod_cleaner)

# Visualize new model
ggplot(data = nontrivial_players, aes(x = OBP, y = SLG)) +
  geom_point() + 
  geom_smooth(method = "lm")
```
  
# High leverage points
Not all points of high leverage are influential. While the high leverage observation corresponding to Bobby Scales in the previous exercise is influential, the three observations for players with OBP and SLG values of 0 are not influential.

This is because they happen to lie right near the regression anyway. Thus, while their extremely low OBP gives them the power to exert influence over the slope of the regression line, their low SLG prevents them from using it.

Instructions

The linear model, mod, is available in your workspace. Use a combination of augment(), arrange() with two arguments, and head() to find the top 6 observations with the highest leverage but the 
lowest Cook's distance.

```{r}
# Rank high leverage points
mod %>% augment() %>% arrange(desc(.hat), .cooksd) %>% head()
```

