---
title: "Datacamp_Cluster Analysis in R__Calculating distance between observations"
author: "dizhen"
date: "2019Äê4ÔÂ19ÈÕ"
output: html_document
---
# Calculate & plot the distance between two players
You've obtained the coordinates relative to the center of the field for two players in a soccer match and would like to calculate the distance between them.

In this exercise you will plot the positions of the 2 players and manually calculate the distance between them by using the Euclidean distance formula.

Instructions
Plot their positions from the two_players data frame using ggplot.
Extract the positions of the players into two data frames player1 and player2.
Calculate the distance between player1 and player2 by using the Euclidean distance formula
¡Ì(x1???x2)2+(y1???y2)2

```{r}
# Plot the positions of the players
ggplot(two_players, aes(x = x, y = y)) + 
  geom_point() +
  # Assuming a 40x60 field
  lims(x = c(-30,30), y = c(-20, 20))

# Split the players data frame into two observations
player1 <- two_players[1, ]
player2 <- two_players[2, ]

# Calculate and print their distance using the Euclidean Distance formula
player_distance <- sqrt( (player1$x - player2$x)^2 + (player1$y - player2$y)^2 )
player_distance
```

# Using the dist() function
Using the Euclidean formula manually may be practical for 2 observations but can get more complicated rather quickly when measuring the distance between many observations.

The dist() function simplifies this process by calculating distances between our observations (rows) using their features (columns). In this case the observations are the player positions and the dimensions are their x and y coordinates.

Note: The default distance calculation for the dist() function is Euclidean distance

Instructions

Calculate the distance between two players using the dist() function for the data frame two_players
Calculate the distance between three players for the data frame three_players

```{r}
# Calculate the Distance Between two_players
dist_two_players <- dist(two_players)
dist_two_players

# Calculate the Distance Between three_players
dist_three_players <- dist(three_players)
dist_three_players
```
# Effects of scale
You have learned that when a variable is on a larger scale than other variables in your data it may disproportionately influence the resulting distance calculated between your observations. Lets see this in action by observing a sample of data from the trees data set.

You will leverage the scale() function which by default centers & scales our column features.

Our variables are the following:

Girth - tree diameter in inches
Height - tree height in inches

Instructions

Calculate the distance matrix for the data frame three_trees and store it as dist_trees
Create a new variable scaled_three_trees where the three_trees data is centered & scaled
Calculate and print the distance matrix for scaled_three_trees and store this as dist_scaled_trees
Output both dist_trees and dist_scaled_trees matrices and observe the change of which observations have the smallest distance between the two matrices (hint: they have changed)

```{r}
# Calculate distance for three_trees 
dist_trees <- dist(three_trees)

# Scale three trees & calculate the distance  
scaled_three_trees <- scale(three_trees)
dist_scaled_trees <- dist(scaled_three_trees)

# Output the results of both Matrices
print('Without Scaling')
dist_trees

print('With Scaling')
dist_scaled_trees
```

# Calculating distance between categorical variables
In this exercise you will explore how to calculate binary (Jaccard) distances. In order to calculate distances we will first have to dummify our categories using the dummy.data.frame() from the library dummies

You will use a small collection of survey observations stored in the data frame job_survey with the following columns:

job_satisfaction Possible options: "Hi", "Mid", "Low"
is_happy Possible options: "Yes", "No"

Instructions

Create a dummified data frame dummy_survey
Generate a Jaccard distance matrix for the dummified survey data dist_survey using the dist() function using the parameter method = 'binary'
Print the original data and the distance matrix
Note the observations with a distance of 0 in the original data (1, 2, and 4)

```{r}
# Dummify the Survey Data
dummy_survey <- dummy.data.frame(job_survey)

# Calculate the Distance
dist_survey <- dist(dummy_survey, method = "binary")

# Print the Original Data
job_survey

# Print the Distance Matrix
dist_survey
```

