---
title: 'Datacamp_Cluster Analysis in R__Case Study: National Occupational mean wage'
author: "dizhen"
date: "2019Äê4ÔÂ23ÈÕ"
output: html_document
---
# Hierarchical clustering: Occupation trees
In the previous exercise you have learned that the oes data is ready for hierarchical clustering without any preprocessing steps necessary. In this exercise you will take the necessary steps to build a dendrogram of occupations based on their yearly average salaries and propose clusters using a height of 100,000.

Instructions

Calculate the Euclidean distance between the occupations and store this in dist_oes
Run hierarchical clustering using average linkage and store in hc_oes
Create a denrogram object dend_oes from your hclust result using the function as.dendrogram()
Plot the dendrogram
Using the color_branches() function create & plot a new dendrogram with clusters colored by a cut height of 100,000

```{r}
# Calculate Euclidean distance between the occupations
dist_oes <- dist(oes, method = "euclidean")

# Generate an average linkage analysis 
hc_oes <- hclust(dist_oes, method = "average")

# Create a dendrogram object from the hclust variable
dend_oes <- as.dendrogram(hc_oes)

# Plot the dendrogram
plot(dend_oes)

# Color branches by cluster formed from the cut at a height of 100000
dend_colored <- color_branches(dend_oes, h = 100000)

# Plot the colored dendrogram
plot(dend_oes)
```

# Hierarchical clustering: Preparing for exploration
You have now created a potential clustering for the oes data, before you can explore these clusters with ggplot2 you will need to process the oes data matrix into a tidy data frame with each occupation assigned its cluster.


Instructions

Create the df_oes data frame from the oes data.matrix, making sure to store the rowname as a column (use rownames_to_column() from the tibble library)
Build the cluster assignment vector cut_oes using cutree() with a h = 100,000
Append the cluster assignments as a column cluster to the df_oes data frame and save the results to a new data frame called clust_oes
Use the gather() function from the tidyr() library to reshape the data into a format amenable for ggplot2 analysis and save the tidied data frame as gather_oes

```{r}
dist_oes <- dist(oes, method = 'euclidean')
hc_oes <- hclust(dist_oes, method = 'average')

library(tibble)
library(tidyr)

# Use rownames_to_column to move the rownames into a column of the data frame
df_oes <- rownames_to_column(as.data.frame(oes), var = 'occupation')

# Create a cluster assignment vector at h = 100,000
cut_oes <- cutree(hc_oes, h = 100000)

# Generate the segmented the oes data frame
clust_oes <- mutate(df_oes, cluster = cut_oes)

# Create a tidy data frame by gathering the year and values into two columns
gathered_oes <- gather(data = clust_oes, 
                       key = year, 
                       value = mean_salary, 
                       -occupation, -cluster)

```

# Hierarchical clustering: Plotting occupational clusters
You have succesfully created all the parts necessary to explore the results of this hierarchical clustering work. In this exercise you will leverage the named assignment vector cut_oes and the tidy data frame gathered_oes to analyze the resulting clusters.

Instructions

View the assignments of each occupation to their clustering by sorting the cut_oes vector using sort()
Use ggplot2 to plot each occupation's average income by year and color the lines by the occupation's assigned cluster.

```{r}
# View the clustering assignments by sorting the cluster assignment vector
sort(cut_oes)

# Plot the relationship between mean_salary and year and color the lines by the assigned cluster
ggplot(gathered_oes, aes(x = year, y = mean_salary, color = factor(cluster))) + 
    geom_line(aes(group = occupation))

```
# K-means: Elbow analysis
In the previous exercises you used the dendrogram to propose a clustering that generated 3 trees. In this exercise you will leverage the k-means elbow plot to propose the "best" number of clusters.

Instructions

Use map_dbl() to run kmeans() using the oes data for k values ranging from 1 to 10 and extract the total within-cluster sum of squares value from each model: model$tot.withinss
Store the resulting vector as tot_withinss
Build a new data frame elbow_df containing the values of k and the vector of total within-cluster sum of squares
Use the values in elbow_df to plot a line plot showing the relationship between k and total within-cluster sum of squares

```{r}
# Use map_dbl to run many models with varying value of k (centers)
tot_withinss <- map_dbl(1:10,  function(k){
  model <- kmeans(x = oes, centers = k)
  model$tot.withinss
})

# Generate a data frame containing both k and tot_withinss
elbow_df <- data.frame(
  k = 1:10,
  tot_withinss = tot_withinss
)

# Plot the elbow plot
ggplot(elbow_df, aes(x = k, y = tot_withinss)) +
  geom_line() +
  scale_x_continuous(breaks = 1:10)
```
# K-means: Average Silhouette Widths
So hierarchical clustering resulting in 3 clusters and the elbow method suggests 2. In this exercise use average silhouette widths to explore what the "best" value of k should be.

Instructions

Use map_dbl() to run pam() using the oes data for k values ranging from 2 to 10 and extract the average silhouette width value from each model: model$silinfo$avg.width Store the resulting vector as sil_width
Build a new data frame sil_df containing the values of k and the vector of average silhouette widths
Use the values in sil_df to plot a line plot showing the relationship between k and average silhouette width

```{r}

# Use map_dbl to run many models with varying value of k
sil_width <- map_dbl(2:10,  function(k){
  model <- pam(oes, k = k)
  model$silinfo$avg.width
})

# Generate a data frame containing both k and sil_width
sil_df <- data.frame(
  k = 2:10,
  sil_width = sil_width
)

# Plot the relationship between k and sil_width
ggplot(sil_df, aes(x = k, y = sil_width)) +
  geom_line() +
  scale_x_continuous(breaks = 2:10)
```

