---
title: "Datacamp_Cluster Analysis in R__Hierarchical clustering"
author: "dizhen"
date: "2019Äê4ÔÂ19ÈÕ"
output: html_document
---
# Calculating linkage
Let us revisit the example with three players on a field. The distance matrix between these three players is shown below and is available as the variable dist_players.

From this we can tell that the first group that forms is between players 1 & 2, since they are the closest to one another with a Euclidean distance value of 11.

Now you want to apply the three linkage methods you have learned to determine what the distance of this group is to player 3.

Instructions

Calculate the distance from player 3 to the group of players 1 & 2 using the following three linkage methods.
Complete: the resulting distance is based on the maximum.
Single: the resulting distance is based on the minimum.
Average: the resulting distance is based on the average.

```{r}
# Extract the pair distances
distance_1_2 <- dist_players[1]
distance_1_3 <- dist_players[2]
distance_2_3 <- dist_players[3]

# Calculate the complete distance between group 1-2 and 3
complete <- max(c(distance_1_3, distance_2_3))
complete

# Calculate the single distance between group 1-2 and 3
single <- min(c(distance_1_3, distance_2_3))
single

# Calculate the average distance between group 1-2 and 3
average <- mean(c(distance_1_3, distance_2_3))
average
```


# Assign cluster membership
In this exercise you will leverage the hclust() function to calculate the iterative linkage steps and you will use the cutree() function to extract the cluster assignments for the desired number (k) of clusters.

You are given the positions of 12 players at the start of a 6v6 soccer match. This is stored in the lineup data frame.

You know that this match has two teams (k = 2), let's use the clustering methods you learned to assign which team each player belongs in based on their position.

Notes:

The linkage method can be passed via the method parameter: hclust(distance_matrix, method = "complete")
Remember that in soccer opposing teams start on their half of the field.
Because these positions are measured using the same scale we do not need to re-scale our data.

Instructions

Calculate the Euclidean distance matrix dist_players among all twelve players
Perform the complete linkage calculation for hierarchical clustering using hclust and store this as hc_players
Build the cluster assignment vector clusters_k2 using cutree() with a k = 2
Append the cluster assignments as a column cluster to the lineup data frame and save the results to a new data frame called lineup_k2_complete

```{r}
# Calculate the Distance
dist_players <- dist(lineup)

# Perform the hierarchical clustering using the complete linkage
hc_players <- hclust(dist_players, method = "complete")

# Calculate the assignment vector with a k of 2
clusters_k2 <- cutree(hc_players, k = 2)

# Create a new data frame storing these results
lineup_k2_complete <- mutate(lineup, cluster = clusters_k2)
```

# Exploring the clusters
Because clustering analysis is always in part qualitative, it is incredibly important to have the necessary tools to explore the results of the clustering.

In this exercise you will explore that data frame you created in the previous exercise lineup_k2_complete.

Reminder: The lineup_k2_complete data frame contains the x & y positions of 12 players at the start of a 6v6 soccer game to which you have added clustering assignments based on the following parameters:

Distance: Euclidean
Number of Clusters (k): 2
Linkage Method: Complete

Instructions

Using count() from dplyr, count the number of players assigned to each cluster.
Using ggplot(), plot the positions of the players and color them by cluster assignment.

```{r}
# Count the cluster assignments
count(lineup_k2_complete, cluster)

# Plot the positions of the players and color them using their cluster
ggplot(lineup_k2_complete, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()
```

# Comparing average, single & complete linkage
You are now ready to analyze the clustering results of the lineup dataset using the dendrogram plot. This will give you a new perspective on the effect the decision of the linkage method has on your resulting cluster analysis.

Instructions

Perform the linkage calculation for hierarchical clustering using the linkages: complete, single and average
Plot the three dendrograms side by side and review the changes

```{r}
# Prepare the Distance Matrix
dist_players <- dist(lineup)

# Generate hclust for complete, single & average linkage methods
hc_complete <- hclust(dist_players, method = "complete")
hc_single <- hclust(dist_players, method = "single")
hc_average <- hclust(dist_players, method = "average")

# Plot & Label the 3 Dendrograms Side-by-Side
# Hint: To see these Side-by-Side run the 4 lines together as one command
par(mfrow = c(1,3))
plot(hc_complete, main = 'Complete Linkage')
plot(hc_single, main = 'Single Linkage')
plot(hc_average, main = 'Average Linkage')
```

# Clusters based on height
In previous exercises you have grouped your observations into clusters using a pre-defined number of clusters (k). In this exercise you will leverage the visual representation of the dendrogram in order to group your observations into clusters using a maximum height (h), below which clusters form.

You will work the color_branches() function from the dendextend library in order to visually inspect the clusters that form at any height along the dendrogram.

The hc_players has been carried over from your previous work with the soccer line-up data.

Instructions

Create a dendrogram object dend_players from your hclust result using the function as.dendrogram()
Plot the dendrogram
Using the color_branches() function create & plot a new dendrogram with clusters colored by a cut height of 20
Repeat the above step with a height of 40

```{r}
library(dendextend)
dist_players <- dist(lineup, method = 'euclidean')
hc_players <- hclust(dist_players, method = "complete")

# Create a dendrogram object from the hclust variable
dend_players <- as.dendrogram(hc_players)

# Plot the dendrogram
plot(dend_players)

# Color branches by cluster formed from the cut at a height of 20 & plot
dend_20 <- color_branches(dend_players, h = 20)

# Plot the dendrogram with clusters colored below height 20
plot(dend_20)

# Color branches by cluster formed from the cut at a height of 40 & plot
dend_40 <- color_branches(dend_players, h = 40)

# Plot the dendrogram with clusters colored below height 40
plot(dend_40)
```

# Exploring the branches cut from the tree
The cutree() function you used in exercises 5 & 6 can also be used to cut a tree at a given height by using the h parameter. Take a moment to explore the clusters you have generated from the previous exercises based on the heights 20 & 40.

Instructions

Build the cluster assignment vector clusters_h20 using cutree() with a h = 20
Append the cluster assignments as a column cluster to the lineup data frame and save the results to a new data frame called lineup_h20_complete
Repeat the above two steps for a height of 40, generating the variables clusters_h40 and lineup_h40_complete
Use ggplot2 to create a scatter plot, colored by the cluster assignment for both heights

```{r}

dist_players <- dist(lineup, method = 'euclidean')
hc_players <- hclust(dist_players, method = "complete")

# Calculate the assignment vector with a h of 20
clusters_h20 <- cutree(hc_players, h = 20)

# Create a new data frame storing these results
lineup_h20_complete <- mutate(lineup, cluster = clusters_h20)

# Calculate the assignment vector with a h of 40
clusters_h40 <- cutree(hc_players, h = 40)

# Create a new data frame storing these results
lineup_h40_complete <- mutate(lineup, cluster = clusters_h40)

# Plot the positions of the players and color them using their cluster for height = 20
ggplot(lineup_h20_complete, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()

# Plot the positions of the players and color them using their cluster for height = 40
ggplot(lineup_h40_complete, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()

```


# Segment wholesale customers
You're now ready to use hierarchical clustering to perform market segmentation (i.e. use consumer characteristics to group them into subgroups).

In this exercise you are provided with the amount spent by 45 different clients of a wholesale distributor for the food categories of Milk, Grocery & Frozen. This is stored in the data frame customers_spend. Assign these clients into meaningful clusters.

Note: For this exercise you can assume that because the data is all of the same type (amount spent) and you will not need to scale it.

Instructions

Calculate the Euclidean distance between the customers and store this in dist_customers
Run hierarchical clustering using complete linkage and store in hc_customers
Plot the dendrogram
Create a cluster assignment vector using a height of 15,000 and store it as clust_customers
Generate a new data frame segment_customers by appending the cluster assignment as the column cluster to the original customers_spend data frame

```{r}
# Calculate Euclidean distance between customers
dist_customers <- dist(customers_spend)

# Generate a complete linkage analysis 
hc_customers <- hclust(dist_customers, method = "complete")

# Plot the dendrogram
plot(hc_customers)

# Create a cluster assignment vector at h = 15000
clust_customers <- cutree(hc_customers, h = 15000)

# Generate the segmented customers data frame
segment_customers <- mutate(customers_spend, cluster = clust_customers)
```
# Explore wholesale customer clusters
Continuing your work on the wholesale dataset you are now ready to analyze the characteristics of these clusters.

Since you are working with more than 2 dimensions it would be challenging to visualize a scatter plot of the clusters, instead you will rely on summary statistics to explore these clusters. In this exercise you will analyze the mean amount spent in each cluster for all three categories.

Instructions

Calculate the size of each cluster using count().
Color & plot the dendrogram using the height of 15,000.
Calculate the average spending for each category within each cluster using the summarise_all() function.

```{r}

dist_customers <- dist(customers_spend)
hc_customers <- hclust(dist_customers)
clust_customers <- cutree(hc_customers, h = 15000)
segment_customers <- mutate(customers_spend, cluster = clust_customers)

# Count the number of customers that fall into each cluster
count(segment_customers, cluster)

# Color the dendrogram based on the height cutoff
dend_customers <- as.dendrogram(hc_customers)
dend_colored <- color_branches(dend_customers, h = 15000)

# Plot the colored dendrogram
plot(dend_colored)

# Calculate the mean for each category
segment_customers %>% 
  group_by(cluster) %>% 
  summarise_all(funs(mean(.)))
```

  