---
title: "Interactive Data Visualization with plotly in R_Advanced charts"
author: "dizhen"
date: "2020/4/16"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Layering traces

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(plotly)

wine <- read_csv("data/wine_data/wine.csv")
glimpse(wine)
```

Adding a smoother

```{r}
m <- loess(Alcohol ~ Flavanoids, data = wine, span = 1.5)
wine %>%
  plot_ly(x = ~Flavanoids, y = ~Alcohol) %>%
  add_markers() %>%
  add_lines(y = ~fitted(m)) %>%
  layout(showlegend = FALSE)
```

Adding a second smoother

```{r}
m2 <- lm(Alcohol ~ poly(Flavanoids, 2), data = wine)
wine %>%
  plot_ly(x = ~Flavanoids, y = ~Alcohol) %>%
  add_markers(showlegend = FALSE) %>%
  add_lines(y = ~fitted(m), name = "LOESS") %>%
  add_lines(y = ~fitted(m2), name = "Polynomial")
```

Layering densities

```{r}
d1 <- filter(wine, Type == 1)
d2 <- filter(wine, Type == 2)
d3 <- filter(wine, Type == 3)


density1 <- density(d1$Flavanoids)
density2 <- density(d2$Flavanoids)
density3 <- density(d3$Flavanoids)

plot_ly(opacity = 0.5) %>%
  add_lines(x = ~density1$x, y = ~density1$y, name = "Type 1") %>%
  add_lines(x = ~density2$x, y = ~density2$y, name = "Type 2") %>%
  add_lines(x = ~density3$x, y = ~density3$y, name = "Type 3") %>%
  layout(xaxis = list(title = 'Flavonoids'),yaxis = list(title = 'Density'))
```

### Practice

Adding a linear smoother

```{r}
vgsales <- read_csv("data/vgsales.csv")
vgsales$User_Score <- as.numeric(vgsales$User_Score)
vgsales$Critic_Score <- as.numeric(vgsales$Critic_Score)
vgsales2016 <- vgsales %>% filter(Year == 2016) %>% na.omit()


# Fit the regression model of User_Score on Critic_Score
m <- lm(User_Score ~ Critic_Score, data = vgsales2016)

# Create the scatterplot with smoother
vgsales2016 %>%
   select(User_Score, Critic_Score) %>%
   na.omit() %>%
   plot_ly(x = ~Critic_Score, y = ~User_Score) %>%
   add_markers(showlegend = FALSE) %>%
   add_lines(y = ~fitted(m))
```

Overlayed density plots

```{r}
d <- density(vgsales2016$Critic_Score, na.rm = TRUE)
plot_ly() %>%
  add_lines(x = ~d$x, y = ~d$y, fill = 'tozeroy') %>%
  layout(xaxis = list(title = 'Critic score'),
         yaxis = list(title = 'Density'))
```


```{r}
# Compute density curves
d.a <- density(filter(vgsales, Publisher== "Activision")$Critic_Score, na.rm = TRUE)
# d.e <- density(filter(vgsales, Publisher== "EA")$Critic_Score, na.rm = TRUE)
d.n <- density(filter(vgsales, Publisher== "Nintendo")$Critic_Score, na.rm = TRUE)

# Overlay density plots
plot_ly() %>%
  add_lines(x = ~d.a$x, y = ~d.a$y, name = "Activision", fill = 'tozeroy') %>%
  # add_lines(x = ~d.e$x, y = ~d.e$y, name = "Electronic Arts", fill = 'tozeroy') %>%
  add_lines(x = ~d.n$x, y = ~d.n$y, name = "Nintendo", fill = 'tozeroy') %>%
  layout(xaxis = list(title = 'Critic Score'),
         yaxis = list(title = 'Density'))

```

## Subplots

1. Representing many categories

```{r}
vgsales2016 %>%
  plot_ly(x = ~Critic_Score, y = ~User_Score, color = ~Genre) %>%
  add_markers()
```

2. A single subplot

```{r}
action_df <- vgsales2016 %>%
  filter(Genre == "Action")
glimpse(action_df)

action_df %>%
  plot_ly(x = ~Critic_Score, y = ~User_Score) %>%
  add_markers()
```
3. Two subplots

```{r}
p1 <- action_df %>%
  plot_ly(x = ~Critic_Score, y = ~User_Score) %>%
  add_markers()
p2 <- vgsales2016 %>%
  filter(Genre == "Adventure") %>%
  plot_ly(x = ~Critic_Score, y = ~User_Score) %>%
  add_markers()

subplot(p1, p2, nrows = 1)
```

4. Legends

```{r}
p1 <- action_df %>%
  plot_ly(x = ~Critic_Score, y = ~User_Score) %>%
  add_markers(name = ~Genre)
p2 <- vgsales2016 %>%
  filter(Genre == "Adventure") %>%
  plot_ly(x = ~Critic_Score, y = ~User_Score) %>%
  add_markers(name = ~Genre)

subplot(p1, p2, nrows = 1)
```

5. Axis labels

```{r}
subplot(p1, p2, nrows = 1, shareY = TRUE, shareX = TRUE)
```


6. group_by() + do()
```{r}
vgsales2016 %>%
  group_by(Genre) %>%
  do(plot = plot_ly(data = ., x = ~Critic_Score, y = ~User_Score) %>%
  add_markers(name = ~Genre)) %>%
  subplot(nrows = 2)
```

### Practice

Manual faceting


```{r}
# Create a scatterplot of User_Score against Critic_Score for PS4 games
p1 <- vgsales2016 %>%
   filter(Platform == "PS4") %>%
   plot_ly(x = ~Critic_Score, y = ~User_Score) %>% 
   add_markers(name = "PS4")

# Create a scatterplot of User_Score against Critic_Score for XOne games
p2 <- vgsales2016 %>%
   filter(Platform == "XOne") %>%
   plot_ly(x = ~Critic_Score, y = ~User_Score) %>% 
   add_markers(name = "XOne")

# Create a facted scatterplot containing p1 and p2
subplot(p1, p2, nrows = 2)
```

Automated faceting

```{r}
# Create a faceted scatterplot of User_Score vs. Critic_Score with 3 rows
vgsales2016 %>%
  group_by(Platform) %>%
  do(
    plot = plot_ly(data = ., x=~Critic_Score, y=~User_Score) %>%
      add_markers(name = ~Platform)
  ) %>%
  subplot(nrows = 3, shareY = TRUE, shareX = TRUE)
```

By default, the subplot() command sets titleX = shareX and titleY = shareY; thus, axis labels are only displayed if shareX and/or shareY are TRUE. You can add titleX = TRUE and/or titleY = TRUE to override this behavior.

```{r}
# Add x-axis and y-axis labels, and a title
subplot(p1, p2, nrows = 2, shareX = TRUE, shareY = TRUE) %>%
   layout(title = "User score vs. critic score by platform, 2016")
```
multiple x-axes to rename

```{r eval=FALSE}
p %>% # subplot
   layout(
     xaxis = list(title = "title 1"), 
     xaxis2 = list(title = "title 2"),
     xaxis3 = list(title = "title 3"), 
     xaxis4 = list(title = "title 4")
   )
```
```{r}
library(forcats)

a <- vgsales%>%
  filter(Publisher == "Activision"| Publisher == "Sony Computer Entertainment" |Publisher == "Ubisoft" | Publisher == "Electronic Arts"| Publisher == "Nintendo" ) %>%
  group_by(Publisher)%>%
  summarise(total_sales = sum(Global_Sales))%>%
  mutate(Publisher = fct_reorder(factor(Publisher), total_sales, .desc = TRUE))

s1<- a%>%plot_ly(x=~Publisher,y=~total_sales, color = ~Publisher) %>% add_bars(name = ~Publisher)
s1

  
b <- vgsales%>%
  filter(Publisher == "Activision"| Publisher == "Sony Computer Entertainment" |Publisher == "Ubisoft" | Publisher == "Electronic Arts"| Publisher == "Nintendo" ) %>%
  group_by(Publisher,Year)%>%
  summarise(total_sales = sum(Global_Sales))

s2 <- b%>%plot_ly(x=~Year,y=~total_sales, color = ~Publisher) %>% add_lines(name = ~Publisher)

s2

sp2 <- subplot(s1, s2, nrows = 1, shareX = FALSE, shareY = FALSE) 
sp2

# Add x-axis and y-axis labels, and a title to  sp2
sp2 %>%
   layout(
     xaxis = list(title = ""),
     xaxis2 = list(title = "Year"),
     yaxis = list(title = "Global Sales (M units)"),
     yaxis2 = list(title = "Global Sales (M units)")
   )
```

## Scatterplot matrices

1. The template

* add_trace() to specify variables

* For each variable,two arguments:

String for axis label

Mapping specifying variable

* 'splom' trace type = scatterplot matrix

```{r eval=FALSE}
data %>%
  plot_ly() %>%
  add_trace(
    type = 'splom',
    dimensions = list(
      list(label='string-1', values=X1),
      list(label='string-2', values=X2),
      .
      .
      .
      list(label='string-n', values=Xn))
  )
```

2. Wine SPLOM

```{r}
wine %>%
  plot_ly() %>%
  add_trace(type = 'splom',
            dimensions = list(
              list(label='Alcohol', values=~Alcohol),
              list(label='Flavonoids', values=~Flavanoids),
              list(label='Color', values=~Color))
            )


```

3. Adding color

```{r}
wine %>%
  plot_ly(color = ~factor(Type)) %>%
  add_trace( type = 'splom',
             dimensions = list(
               list(label='Alcohol', values=~Alcohol),
               list(label='Flavonoids', values=~Flavanoids),
               list(label='Color', values=~Color))
             )
```

### Practice

```{r}
# Create a SPLOM of NA_Sales, EU_Sales, and JP_Sales
vgsales2016 %>%
  plot_ly() %>%
  add_trace(
    type = 'splom',
    dimensions = list(
      list(label = 'N. America', values = ~NA_Sales),
      list(label = 'Europe', values = ~EU_Sales),
      list(label = 'Japan', values = ~JP_Sales)
    )
  )
```

```{r}

# Color the SPLOM of NA_Sales, EU_Sales, and JP_Sales by nintendo
splom <- vgsales2016 %>%
  mutate(nintendo = ifelse(Publisher == "Nintendo", "Nintendo", "Other")) %>%
  plot_ly(color = ~nintendo) %>% 
  add_trace(
    type = "splom",                                  
    dimensions = list(
      list(label = 'N. America', values = ~NA_Sales),
      list(label = 'Europe', values = ~EU_Sales),
      list(label = 'Japan', values = ~JP_Sales
    )
  ))
  
splom
```

```{r}
# Delete the diagonal plots in splom
splom %>%
   style(diagonal = list(visible = FALSE))

# Delete the plots in the upper half of splom
splom %>%
   style(showupperhalf = FALSE)

# Delete the plots in the lower half of splom
splom %>%
   style(showlowerhalf = FALSE)

```


## Binned scatterplots

```{r eval = FALSE}
sim_data %>%
  plot_ly(x = ~x, y = ~y) %>%
  add_histogram2d()

sim_data %>%
  plot_ly(x = ~x, y = ~y) %>%
  add_histogram2d(nbinsx = 200, nbinsy = 100)

```

### Practice

```{r}
# Create a binned scatterplot of User_Score vs. Critic_Score
vgsales %>%
  plot_ly(y=~User_Score, x=~Critic_Score) %>%
  add_histogram2d(nbinsy=50, nbinsx=50)
```