---
title: "Interactive Data Visualization with plotly in R_Case Study"
author: "dizhen"
date: "2020/4/18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(plotly)
library(forcats)

turnout <- read_csv("data/election_data/TurnoutRates.csv")
fundraising <- read_csv("data/election_data/fec_candidate_summary_2018.csv")
senate_winners <- read_csv("data/election_data/senate_winners.csv")
fl_boundaries <- read_csv("data/election_data/fl_boundaries.csv")
fl_results <- read_csv("data/election_data/fl_results.csv")
```

## Introduction to the 2018 election data

```{r}
# Create a scatterplot of turnout2018 against turnout2014
p <- turnout %>%
  plot_ly(x = ~turnout2014, y=~turnout2018) %>%
  add_markers() %>%
  layout(xaxis = list(title = "2014 voter turnout"),
         yaxis = list(title = "2018 voter turnout"))
         
p
```
You can add a line by connecting two points using add_lines(x = c(x1, x2), y = (y1, y2))

```{r}
# Add the line y = x to the scatterplot
p %>%
  add_lines(x = c(0.25, 0.6), y = c(0.25, 0.6)) %>%
  layout(showlegend = FALSE)
```
In the sample code, turnout %>% top_n(15, wt = turnout) extracts the 15 states with the highest turnout rates.

```{r}

# Create a dotplot of voter turnout in 2018 by state ordered by turnout
turnout %>%
  top_n(15, wt = turnout2018) %>%
  plot_ly(x = ~turnout2018, y = ~fct_reorder(state, turnout2018)) %>%
  add_markers() %>%
  layout(xaxis = list(title = "Eligible voter turnout"), 
         yaxis = list(title = "State", type= "category"))
```

```{r}
# Create a histogram of receipts for the senate races
fundraising %>%
  filter(office == "S") %>%
  plot_ly(x = ~receipts) %>%
  add_histogram() %>%
  layout(title = "Fundraising for 2018 Senate races",
         xaxis = list(title = "Total contributions received"))
```

```{r}
# Create a dotplot of the top 15 Senate campaigns
fundraising %>%
  filter(office == "S") %>%
  top_n(15, wt = receipts) %>%
  plot_ly(x = ~receipts, y = ~fct_reorder(state, receipts),
          color = ~fct_drop(party),
          hoverinfo = "text",
          text = ~paste("Candidate:", name, "<br>",
                        "Party:", party, "<br>",
                        "Receipts:", receipts, "<br>",
                        "Disbursements:", disbursement)) %>%
  add_markers(colors = c("blue", "red")) 
```


## Choropleth maps

1. Choropleth maps in plotly


```{r eval=FALSE}
turnout %>%
  plot_geo(locationmode = 'USA-states') %>%
  add_trace(
    z = ~turnout, # Sets the color values
    locations = ~state.abbr # Matches cases to polygons
    ) %>%
  layout(geo = list(scope = 'usa')) # Restricts map only to USA
```

2. Limitation of plot_geo()

locationmode: "USA-states" | "ISO-3" | "country names"

3. Mapping options

* scope = "usa" 

"world"| "usa"| "europe"| "asia"| "africa"| "north america"| "south america"

* projection = list(type ="mercator")

"conic conformal"| "mercator"| "robinson"| "stereographic"| and 18 more...

* scale = 1

Larger values = tighter zoom

* center = list(lat = ~c.lat, lon = ~c.lon)

Set c.lat and c.lon to center the map

```{r}
# Create a choropleth map of the change in voter turnout from 2014 to 2018
turnout %>%
  mutate(change = turnout2018 - turnout2014) %>%
  plot_geo(locationmode = 'USA-states') %>%
  add_trace(z = ~change, locations = ~state.abbr) %>%
  layout(geo = list(scope = "usa"))
```

This task requires you to map a factor to the fill color. However, the z aesthetic expects a numeric variable. An easy work around is to convert party to a numeric variable via as.numeric(party) and then manually specify the desired colors in add_trace(). Additionally, the colorbar is no longer very useful, and can be removed by adding the layer hide_colorbar().

```{r}
# Create a choropleth map displaying the Senate results
senate_winners %>%
  plot_geo(locationmode = 'USA-states') %>%
  add_trace(z = ~as.numeric(party), locations = ~state,
    colors = c("dodgerblue", "mediumseagreen", "tomato"),
    hoverinfo = "text",
    text = ~paste("Candidate:", name, "<br>",
                  "Party:", party, "<br>",
                  "% vote:", round(pct.vote, 1))
  ) %>%
  layout(geo = list(scope = 'usa')) %>% 
  hide_colorbar()
```
```{r eval=FALSE}

# Map President Trump's rallies in 2018
rallies2018 %>%
  plot_geo(locationmode = 'USA-states') %>%
  add_markers(
    x = ~long, y = ~lat, size = ~no.speakers,  
    hoverinfo = "text", text = ~paste(city, state, sep = ",")
  ) %>%
  layout(title = "2018 Trump Rallies", 
      geo = list(scope = "usa"))
```
This it is quite easy to customize by specifying additional arguments in the list passed to geo in layout().

In this exercise you will explore a few useful options outlined below:

* To change the color of the landmass, add the arguments showland = TRUE and set a landcolor.

* To make lakes distinct from landmasses, add the arguments showlakes = TRUE and set a lakecolor.

* To display states/provinces, set showsubunit = TRUE, and the set subunitcolor.

* To display countries, set showcountries = TRUE, and the set countrycolor.

Note that you must use the toRGB() function in order to pass R colors to the geo layout.

```{r eval = FALSE}
# Customize the geo layout
g <- list(scope = 'usa', 
          showland = TRUE, landcolor = toRGB("gray90"),
          showlakes = TRUE, lakecolor = toRGB("white"),
          showsubunit = TRUE, subunitcolor = toRGB("white"))

# Apply the geo layout to the map
rallies2018 %>%
  plot_geo(locationmode = 'USA-states') %>%
  add_markers(
    x = ~long, y = ~lat, size = ~no.speakers, 
    hoverinfo = "text", text = ~paste(city, state, sep = ",")
  ) %>%
  layout(title = "2018 Trump Rallies", geo = g)
```


## From polygons to maps

1. Joining data frames

```{r eval = FALSE}
turnout <- turnout %>%
  mutate(state = tolower(state)) # make state names lowercase
  
states_map <- left_join(us_states, turnout, by = c("region" = "state"))
```

2. Creating the map

```{r eval = FALSE}
states_map %>%
  group_by(group) %>%
  plot_ly(
    x = ~long,
    y = ~lat,
    color = ~turnout2018, # variable mapped to fill color
    split = ~region # no more than one fill color per polygon
  ) %>%
  add_polygons(
    line = list(width = 0.4),
    showlegend = FALSE
  )
```


3. Polishing your map

```{r eval=FALSE}
state_turnout_map %>%
  layout(
    title = "2018 Voter Turnout by State", 
    xaxis = list(title ="", showgrid = FALSE,
                 zeroline = FALSE, showticklabels = FALSE),
    yaxis = list(title ="", showgrid = FALSE,
                 zeroline = FALSE, showticklabels = FALSE)
  ) %>%
  colorbar(title ="Turnout")
```


```{r eval=FALSE}
# Create a choropleth map displaying the Senate winners
senate_map %>%
  group_by(group) %>%
  plot_ly(x=~long, y=~lat, color = ~party, split = ~region) %>%
  add_polygons(line = list(width = 0.4), showlegend=FALSE)

# Adjust the polygon colors and boundaries
senate_map %>%
  group_by(group) %>%
  plot_ly(x = ~long, y = ~lat, color = ~party, split = ~region,
          colors = c("dodgerblue", "mediumseagreen", "tomato")) %>%
  add_polygons(line = list(width = 0.4,color = toRGB("gray60")), showlegend = FALSE)

# Define the layout settings to polish the axes
map_axes <- list(title = "", showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)

# Apply the layout to both axes
senate_map %>%
  group_by(group) %>%
  plot_ly(x = ~long, y = ~lat, color = ~party, split = ~region,
          colors = c("dodgerblue", "mediumseagreen", "tomato")) %>%
  add_polygons(line = list(width = 0.4, color = toRGB("gray60")), showlegend = FALSE) %>%
  layout(xaxis = map_axes, yaxis = map_axes)


```

```{r}
# Join the fl_boundaries and fl_results data frames
senate_vote <- left_join(fl_boundaries, fl_results, by = c("subregion" = "CountyName"))

# Create a county-level choropleth map of Pctvote
senate_vote %>%
  group_by(group) %>%
  plot_ly(x = ~long, y = ~lat, 
          color = ~Pctvote,
          split = ~subregion) %>%
  add_polygons(line = list(width = 0.4), showlegend = FALSE, colors = c("blue", "red"))

# Join the fl_boundaries and fl_results data frames
senate_vote <- left_join(fl_boundaries, fl_results, by = c("subregion" = "CountyName"))

# Specify the axis settings to polish the map
map_axes <- list(title="", showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)

# Create a polished county-level choropleth map of Pctvote
senate_vote %>%
  group_by(group) %>%
  plot_ly(x = ~long, y = ~lat, 
          color = ~Pctvote,
          split = ~subregion) %>%
  add_polygons(line = list(width = 0.4), showlegend = FALSE, colors = c("blue", "red")) %>%
  layout(xaxis = map_axes, yaxis = map_axes)
```


## Going further

Books

* [Plotly for R](https://plotly-r.com/) by Carson Sievert

* [Interactive Graphics for Data Analysis](https://www.crcpress.com/Interactive-Graphics-for-Data-Analysis-Principles-and-Examples/Theus-Urbanek/p/book/9781584885948)

* [Interactive and Dynamic Graphics for Data Analysis](https://link.springer.com/book/10.1007/978-0-387-71762-3)

Community

* [community.plot.ly](https://community.plotly.com/) don't hesitate to ask the community a question!

Manuals

* [plot.ly/r](https://plotly.com/r/): Quick examples along with indepth documentation

* [plotly cheatsheet](https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf?_ga=2.262472844.1660058896.1542557393-851051236.1530303515)

Forging ahead

* [Putting your charts online](https://plotly.com/r/getting-started/#initialization-for-online-plotting)

* Linked and/or animated graphics

* shiny + plot.ly