---
title: "Interactive Data Visualization with plotly in R_Styling and customizing your graphics"
author: "dizhen"
date: "2020/4/16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Customize your traces

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(plotly)

wine <- read_csv("data/wine_data/wine.csv")
glimpse(wine)
winequality <- read_csv("data/wine_data/winequality.csv")
glimpse(winequality)
vgsales <- read_csv("data/vgsales.csv")
glimpse(vgsales)
```

1. Color

```{r}
winequality %>%
  plot_ly(x = ~fixed_acidity) %>%
  add_histogram()

winequality %>%
  plot_ly(x = ~fixed_acidity) %>%
  add_histogram(color = I("red")) # Setting color

```

2. Opacity

```{r}
winequality %>%
  plot_ly(x = ~residual_sugar, y = ~fixed_acidity) %>%
  add_markers()


winequality %>%
  plot_ly(x = ~residual_sugar, y = ~fixed_acidity) %>%
  add_markers(marker = list(opacity = 0.2)) # Adjust opacity
```

3. Symbols

```{r}
winequality %>%
  plot_ly(x = ~residual_sugar, y = ~fixed_acidity) %>%
  add_markers(marker = list(symbol = "circle-open")) # Change symbol
```

4.Marker options

* opacity

* color

* symbol (scatter/box)

* size (scatter)

* width (bar/histogram)


### Practice

```{r}
vgsales2016 = vgsales %>%
	filter(Year == 2016)

# Create a histogram of Critic_Score with navy bars that are 50% transparent
vgsales2016 %>%
  plot_ly(x = ~Critic_Score) %>%
  add_histogram(color = I("navy"), opacity = 0.5)
```

```{r}
# Change the color of the histogram using a hex code
vgsales2016 %>%
  plot_ly(x = ~Critic_Score) %>%
  add_histogram(color = I("#111e6c"))

# Change the color of the histogram using rgb()
vgsales2016 %>%
  plot_ly(x = ~Critic_Score) %>%
  add_histogram(marker = list(color = "rgb(17, 30, 108)"))
```

```{r}
# Set the plotting symbol to diamond and the size to 4
plot_ly(data = vgsales2016, x = ~User_Score, y = ~Critic_Score) %>% 
	add_markers(marker = list(symbol = "diamond", size=4)) 
```

## Thoughtful use of color

1. Adding a third variable
```{r}
wine %>%
  plot_ly(x = ~Flavanoids, y = ~Alcohol, color = ~Type) %>%
  add_markers()
```

2. Adding a quantitative variable
```{r}
wine %>%
  plot_ly(x = ~Flavanoids, y = ~Alcohol, color = ~Color) %>%
  add_markers()
```

3. RColorBrewer palettes
```{r}
wine %>%
  plot_ly(x = ~Flavanoids, y = ~Alcohol, color = ~Type) %>%
  add_markers(colors = "Dark2")
```

4. Manual palettes
```{r}
wine %>%
  plot_ly(x = ~Flavanoids, y = ~Alcohol, color = ~Type) %>%
  add_markers(colors = c("orange", "black", "skyblue"))
```

### Practice

```{r}
# Use color to add Genre as a third variable
vgsales2016 %>%
  plot_ly(x=~Critic_Score, y=~User_Score, color = ~Genre) %>%
  add_markers(colors = "Dark2")
```

```{r}
# Create a scatterplot of User_Score against Critic_Score coded by Rating
vgsales2016 %>%
   plot_ly(x = ~Critic_Score, y = ~User_Score, symbol = ~Rating) %>%
   add_markers()
```

```{r}
# Create a scatterplot of User_Score vs. Critic_Score colored by User_Count
vgsales2016 %>%
  plot_ly(x = ~Critic_Score, y = ~User_Score, color = ~User_Count) %>%
  add_markers()

```

```{r}
# Create a scatterplot of User_Score vs. Critic_Score colored by log User_Count
vgsales2016 %>%
  plot_ly(x = ~Critic_Score, y = ~User_Score,color = ~log(User_Count)) %>%
  add_markers()
```


## Hover info (Labeling your data)

1. Changing the default

hoverinfo =

* "all"

* "x"

* "y"

* "x+y"

* "x+y+z"


```{r}
wine %>%
  count(Type) %>%
  plot_ly(x = ~Type, y = ~n,
          
          # Changing the default info displaye
          hoverinfo = "y") %>%
  add_bars()
```

2. Custom hover text

tilde,~,to map columns to aesthetic parameters

```{r}
wine %>%
  plot_ly(x = ~Flavanoids, y = ~Alcohol, hoverinfo = "text", 
          text = ~paste("Flavanoids:", Flavanoids, "<br>", "Alcohol:", Alcohol)) %>%
  add_markers()

```


### Practice

```{r}
# Create a bar chart of Platform with hoverinfo only for the bar heights
vgsales2016 %>%
	count(Platform) %>%
	plot_ly(x = ~Platform, y = ~n, hoverinfo = "y") %>%
 	add_bars()
```

```{r}

# Create a scatterplot of User_Score vs. Critic score
vgsales2016 %>%
	# Add video game Name to the hover info text
	plot_ly(x = ~Critic_Score, y = ~User_Score, text = ~Name) %>% 
	add_markers()
```

```{r}

# Format the hover info for NA_Sales, EU_Sales, and Name
vgsales2016 %>%
  plot_ly(x = ~NA_Sales, y = ~EU_Sales,
          hoverinfo = "text",
          text = ~paste("NA_Sales:", NA_Sales, "<br>",
                     "EU_Sales:", EU_Sales, "<br>",
                     "Name:", Name)
  ) %>%
  add_markers()
```


## Customizing your layout

layout()

* axes:type, labels,tick marks,transformations, etc.

* legend: position

* canvas: grid lines, background color

* size: height, width, margins

1. Axis labels

```{r}
winequality %>%
  plot_ly(x = ~free_so2, y = ~total_so2) %>%
  add_markers(marker = list(opacity = 0.2))

winequality %>%
  plot_ly(x = ~free_so2, y = ~total_so2) %>%
  add_markers(marker = list(opacity = 0.2)) %>%
  layout(xaxis = list(title = "Free SO2 (ppm)"), yaxis = list(title = "Total SO2 (ppm)"))

```

2. Titles

```{r}
winequality %>%
  plot_ly(x = ~free_so2, y = ~total_so2) %>%
  add_markers(marker = list(opacity = 0.2)) %>%
  layout(xaxis = list(title = "Free SO2 (ppm)"), 
         yaxis = list(title = "Total SO2 (ppm)"), 
         title = "Does free SO2 predict total SO2 in wine?")
```

3. Transforming axes

```{r}
winequality %>%
  plot_ly(x = ~free_so2, y = ~total_so2) %>%
  add_markers(marker = list(opacity = 0.2)) %>%
  layout(xaxis = list(title = "Free SO2 (ppm, log scale)", type = "log"),
         yaxis = list(title = "Total SO2 (ppm, log scale)", type = "log"),
         title = "Does free SO2 predict total SO2 in wine?")
```

4. Customizing the grid
```{r}
winequality %>%
  plot_ly(x = ~free_so2, y = ~total_so2) %>%
  add_markers(marker = list(opacity = 0.5)) %>%
  layout(xaxis = list(title = "Free SO2 (ppm)" , zeroline = FALSE),
         yaxis = list(title = "Total SO2 (ppm)", zeroline = FALSE, showgrid = FALSE))
```

5. Customizing the canvas 

```{r}
winequality %>%
  plot_ly(x = ~free_so2, y = ~total_so2) %>%
  add_markers(marker = list(opacity = 0.5)) %>%
  layout(xaxis = list(title = "Free SO2 (ppm)"),
         yaxis = list(title = "Total SO2 (ppm)"),
         plot_bgcolor = toRGB("gray90"),
         paper_bgcolor = toRGB("skyblue"))

```

### Practice

```{r}

# Polish the scatterplot by transforming the x-axis and labeling both axes
vgsales2016 %>%
  plot_ly(x = ~Global_Sales, y = ~Critic_Score) %>%
  add_markers(marker = list(opacity = 0.5)) %>%
  layout(xaxis = list(title="Global sales (millions of units)",type = "log"),
         yaxis = list(title = "Critic score"))

```

```{r}
annual_vgsales<- vgsales %>% group_by(Year) %>% summarise(Global_Sales = sum(Global_Sales))

# Set the background color to #ebebeb and remove the vertical grid
annual_vgsales %>%
  plot_ly(x = ~Year, y = ~Global_Sales) %>%
  add_lines() %>%
  layout(xaxis = list(showgrid = FALSE),
  paper_bgcolor = "#ebebeb")
```