---
title: "Datacamp_Cleaning Data in R_Introduction and exploring raw data"
author: "dizhen"
date: "2019/4/4"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to cleaning data in R


```{r}

weather <- readRDS("data/weather.rds")

# View the first 6 rows of data
head(weather)

# View the last 6 rows of data
tail(weather)

# View a condensed summary of the data
str(weather)
```



## Exploring raw data

1. Understanding the structure of your data

class() - Class of data object

dim() - Dimensions of data

names() - Column names

str() - Preview of data with helpful details

glimpse() - Be!er version of str() from dplyr

summary() - Summary of data

2. Looking at your data

head() - View top of dataset

tail() - View bo!om of dataset

print() - View entire dataset (not recommended!)

3. Visualizing your data

hist() - View histogram of a single variable

plot() - View plot of two variables


### Practice

```{r}
library(readr)
bmi <- read_csv("data/bmi_clean.csv")

# Check the class of bmi
class(bmi)

# Check the dimensions of bmi
dim(bmi)

# View the column names of bmi
names(bmi)
```


```{r}
# Check the structure of bmi
str(bmi)

# Load dplyr
library(dplyr)

# Check the structure of bmi, the dplyr way
glimpse(bmi)

# View a summary of bmi
summary(bmi)
```


```{r}
# Print bmi to the console
print(bmi)

# View the first 6 rows
head(bmi,n=6)

# View the first 15 rows
head(bmi,n=15)

# View the last 6 rows
tail(bmi,n = 6)

# View the last 10 rows
tail(bmi, n=10)
```


```{r}
# Histogram of BMIs from 2008
hist(bmi$Y2008)

# Scatter plot comparing BMIs from 1980 to those from 2008
plot(bmi$Y1980,bmi$Y2008)
```

