---
title: "Datacamp_Cleaning Data in R_Preparing data for analysis"
author: "dizhen"
date: "2019/4/5"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Type Conversion

```{r}
as.character(2016) 
as.numeric(TRUE) 
as.integer(99) 
as.factor("something")
as.logical(0)  
```
```{r}
library(lubridate)
# Experiment with basic lubridate functions
ymd("2015-08-25") # year-month-day
ymd("2015 August 25") # year-month-day
mdy("August 25, 2015") # month-day-year
hms("13:33:09") # hour-minute-second
ymd_hms("2015/08/25 13.33.09") # year-month-day hour-minute-second
```

### Practice


```{r}
# Make this evaluate to "character"
class("TRUE")

# Make this evaluate to "numeric"
class(8484.00)

# Make this evaluate to "integer"
class(99L)

# Make this evaluate to "factor"
class(factor("factor"))

# Make this evaluate to "logical"
class(FALSE)
```



```{r}
library(readr)
students <- read_csv("data/students_with_dates.csv")

# Preview students with str()
str(students)

# Coerce Grades to character
students$Grades <- as.character(students$Grades)

# Coerce Medu to factor
students$Medu <- as.factor(students$Medu)

# Coerce Fedu to factor
students$Fedu <- as.factor(students$Fedu)
    
# Look at students once more with str()
str(students)
```

```{r}
# Preview students2 with str()
students2 <- students

# Load the lubridate package
library(lubridate)

# Parse as date
dmy("17 Sep 2015")

# Parse as date and time (with no seconds!)
mdy_hm("July 15, 2012 12:56")

# Coerce dob to a date (with no time)
students2$dob <- ymd(students2$dob)

# Coerce nurse_visit to a date and time
students2$nurse_visit <- ymd_hms(students2$nurse_visit)
    
# Look at students2 once more with str()
str(students2)
```

## String manipulation

1. Key functions in stringr for cleaning data

str_trim() - Trim leading and trailing white space

str_pad() - Pad with additional characters

str_detect() - Detect a pa!ern

str_replace() - Find and replace a pa!ern

```{r}
library("stringr")
# Trim leading and trailing white space
str_trim(" this is a test ") 

# Pad string with zeros
str_pad("24493", width = 7, side = "left", pad = "0") 

# Create character vector of names
friends <- c("Sarah", "Tom", "Alice")

# Search for string in vector
str_detect(friends, "Alice") 

# Replace string in vector
str_replace(friends, "Alice", "David")

```

2. Other helpful functions in base R

tolower() - Make all lowercase

toupper() - Make all uppercase

```{r}
# Make all lowercase 
tolower("I AM TALKING LOUDLY!!") 

# Make all uppercase
toupper("I am whispering...") 
```




```{r}
# Load the stringr package
library(stringr)

# Trim all leading and trailing whitespace
str_trim(c("   Filip ", "Nick  ", " Jonathan"))

# Pad these strings with leading zeros
str_pad(c("23485W", "8823453Q", "994Z"),width = 9, side = 'left', pad = '0')
```



```{r}
# Print state abbreviations
states<- c("al", "ak", "az", "ar", "ca", "co", "ct", "de", "fl", "ga", "hi", "id")

# Make states all uppercase and save result to states_upper
states_upper <- toupper(states)
states_upper

# Make states_upper all lowercase again
tolower(states_upper)
```


```{r}
# Copy of students2: students3
students3 <- students2

# Look at the head of students3
head(students3)

# Detect all dates of birth (dob) in 1997
str_detect(students3$dob,'1997')

# In the sex column, replace "F" with "Female" ...
students3$sex <- str_replace(students3$sex,'F','Female')

# ... and "M" with "Male"
students3$sex <- str_replace(students3$sex,'M','Male')

# View the head of students3
head(students3)
```


## Missing and special values

1. Missing values

May be random, but dangerous to assume

Sometimes associated with variable/outcome of interest

In R, represented as NA

May appear in other forms

\#N/A (Excel), Single dot (SPSS, SAS), Empty string

2. Special values

Inf - "Infinite value" (indicative of outliers?) 

NaN - "Not a number" (rethink a variable?)


3. Finding missing values

```{r}
# Create small dataset
df <- data.frame(A = c(1, NA, 8, NA), B = c(3, NA, 88, 23), C = c(2, 45, 3, 1))

# Check for NAs
is.na(df)

# Are there any NAs?
any(is.na(df)) 

# Count number of NAs
sum(is.na(df))

# Use summary() to find NAs
summary(df) 

```

4. Dealing with missing values

```{r}
# Find rows with no missing values
complete.cases(df)

# Subset data, keeping only complete cases
df[complete.cases(df), ] 

# Another way to remove rows with NAs
na.omit(df) 
```

### Practice

```{r}
social_df <- data.frame(name = factor(c("Sarah", "Tom","David", "Alice")), n_friends = c(244, NA, 145, 43), status = factor(c("Going out!", "","Moving night...", "")))

# Call is.na() on the full social_df to spot all NAs
is.na(social_df)

# Use the any() function to ask whether there are any NAs in the data
any(is.na(social_df))

# View a summary() of the dataset
summary(social_df)

# Call table() on the status column
table(social_df$status)
```



```{r}
# Replace all empty strings in status with NA
social_df$status[social_df$status == ""] <- NA

# Print social_df to the console
social_df

# Use complete.cases() to see which rows have no missing values
complete.cases(social_df)

# Use na.omit() to remove all rows with any missing values
na.omit(social_df)
```


## Outliers and obvious errors

1. outliers


Extreme values distant from other values

Several causes

 -Valid measurements, Variability in measurement, Experimental error, Data entry error

May be discarded or retained depending on cause


```{r}
# Simulate some data 
set.seed(10)
x <- c(rnorm(30, mean = 15, sd = 5), -5, 28, 35)

# View a boxplot
boxplot(x, horizontal = TRUE)
```

2. Finding outliers and errors

```{r}
# Create another small dataset
df2 <- data.frame(A = rnorm(100, 50, 10), B = c(rnorm(99, 50, 10), 500), C = c(rnorm(99, 50, 10), -1)) 
# View a summary
summary(df2) 

# View a histogram
hist(df2$B, breaks = 20)

# View a boxplot
boxplot(df2)
```


### Practice

```{r}
# Look at a summary() of students3
summary(students3)

students3$age <- as.numeric(round((Sys.Date() - students3$dob)/365))

# View a histogram of the age variable
hist(students3$age)

# View a histogram of the absences variable
hist(students3$absences)

# View a histogram of absences, but force zeros to be bucketed to the right of zero
hist(students3$absences, right = FALSE)
```




```{r}
# View a boxplot of age
boxplot(students3$age)

# View a boxplot of absences
boxplot(students3$absences)
```

