---
title: "Datacamp_Data Visualization with ggplot2 (Part 1)__Aesthetics"
author: "dizhen"
date: "2019/4/8"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Visible Aesthetics

Dataframe column mapped onto visible aesthetic.

Aesthetics in aaes(), attributes in geom_().

Typical asethetics:

* x: X axis position

* y: y axis position

* colour: Colour of dots, oulines of other shapes

* fill: fill colour

* size: Diameter of points, thickness of lines

* alpha: Transparency

* linetype: Line dash pattern

* labels: Text on a plot or axes

* shape: Shape of points


### Practice

```{r}
library(ggplot2)

# 1 - Map mpg to x and cyl to y
ggplot(mtcars, aes(x = mpg, y = cyl)) +
  geom_point()
  
# 2 - Reverse: Map cyl to x and mpg to y
ggplot(mtcars, aes(x = cyl, y = mpg)) +
  geom_point()

# 3 - Map wt to x, mpg to y and cyl to col
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point()

# 4 - Change shape and size of the points in the above plot
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point(shape = 1, size = 4)
```

The *color* aesthetic typically changes the outside outline of an object and the fill aesthetic is typically the inside shading. However, as you saw in the last exercise, *geom_point()* is an exception.


```{r}
# am and cyl are factors, wt is numeric
class(mtcars$am)
class(mtcars$cyl)
class(mtcars$wt)

# From the previous exercise
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point(shape = 1, size = 4)

# 1 - Map cyl to fill
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
  geom_point(shape = 1, size = 4)

# 2 - Change shape and alpha of the points in the above plot
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
  geom_point(shape = 21, size = 4, alpha = 0.6)

# 3 - Map am to col in the above plot
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl,col = am)) +
  geom_point(shape = 21, size = 4, alpha = 0.6)
```

Shapes in R can have a value from 1-25. Shapes 1-20 can only accept a *color* aesthetic, but shapes 21-25 have both a *color* and a *fill* aesthetic. See the *pch* argument in *par()* for further discussion.

```{r}
# Map cyl to size
ggplot(mtcars, aes(x = wt, y = mpg, size = cyl)) +
  geom_point()

# Map cyl to alpha
ggplot(mtcars, aes(x = wt, y = mpg, alpha = cyl)) +
  geom_point()

# Map cyl to shape (a continuous variable cannot map to shape )
# ggplot(mtcars, aes(x = wt, y = mpg, shape = cyl)) +
#   geom_point()

# Map cyl to labels
ggplot(mtcars, aes(x = wt, y = mpg, label = cyl)) +
  geom_text()
```

A word about hexadecimal colours: Hexadecimal, literally "related to 16", is a base-16 alphanumeric counting system. Individual values come from the ranges 0-9 and A-F. This means there are 256 possible two-digit values (i.e. 00 - FF). Hexadecimal colours use this system to specify a six-digit code for Red, Green and Blue values ("*#RRGGBB*") of a colour (i.e. Pure blue: "*#0000FF*", black: "*#000000*", white: "*#FFFFFF*"). R can accept hex codes as valid colours.


```{r}
# Define a hexadecimal color
my_color <- "#4ABEFF"

# Draw a scatter plot with color *aesthetic*
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) + geom_point()


# Same, but set color *attribute* in geom layer 
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) + geom_point(color = my_color)


# Set the fill aesthetic; color, size and shape attributes
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) + geom_point(size = 10, shape = 23, color = my_color)

```


```{r}
# Expand to draw points with alpha 0.5
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl))  + geom_point(alpha = 0.5)

  
# Expand to draw points with shape 24 and color yellow
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) + geom_point(shape = 24, color = "yellow")

  
# Expand to draw text with label rownames(mtcars) and color red
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) + geom_text(label = rownames(mtcars), color = "red")
```


```{r}
# Map mpg onto x, qsec onto y and factor(cyl) onto col
ggplot(mtcars, aes(x = mpg, y = qsec, col = factor(cyl)))+ geom_point()

# Add mapping: factor(am) onto shape
ggplot(mtcars, aes(x = mpg, y = qsec, col = factor(cyl), shape = factor(am)))+ geom_point()

# Add mapping: (hp/wt) onto size
ggplot(mtcars, aes(x = mpg, y = qsec, col = factor(cyl), size = (hp/wt),shape = factor(am)))+ geom_point()
```

## Modifying Aesthetics

1. Positions:

* identity

* dodge

* stack

* fill

* jitter

* jitterdodge

2. Scale Functions:

* scale_x... (e.g. scale_x_continuous("Sepal Width", limits = (2,8), breaks = seq(2,8,3), expand = c(8,0)))

* scale_y...

* scale_color...(e.g. scale_color_discrete("Sepal Width", labels = c("Setosa","Versicolour","Virginica")))

* scale_fill...

* scale_shape...

* scale_linetype...

### Practice

```{r}
# The base layer, cyl.am, is available
cyl.am <- ggplot(mtcars, aes(x = factor(cyl), fill = factor(am)))

# Add geom (position = "stack" by default)
cyl.am + 
  geom_bar()

# Fill - show proportion
cyl.am + 
  geom_bar(position = "fill")  

# Dodging - principles of similarity and proximity
cyl.am +
  geom_bar(position = "dodge") 

# Clean up the axes with scale_ functions
val = c("#E41A1C", "#377EB8")
lab = c("Manual", "Automatic")
cyl.am +
  geom_bar(position = "dodge") +
  scale_x_discrete() + 
  scale_y_continuous() +
  scale_fill_manual("Transmission", 
                    values = val,
                    labels = lab) 
```



```{r}
# 1 - Create jittered plot of mtcars, mpg onto x, 0 onto y
ggplot(mtcars, aes(x = mpg, y = 0)) +
  geom_jitter()

# 2 - Add function to change y axis limits
ggplot(mtcars, aes(x = mpg, y = 0)) +
  geom_jitter() + scale_y_continuous(limits = c(-2,2))
```

## Aesthetics Best Practice

1. Which Aesthetics?

* Be creative

* Clear guidelines

* Jacques Bertin -*The Semiology of Graphics, 1967*

* William Cleveland -*Perception of visual elements (90s)*

2. Aesthetics -Categorical Variables

* labels: Text on plot or axes

* fill: Fill color

* shape: Shape of point

* alpha: Transparency

* linetype: Line dash pattern

* size: Diameter of points, thickness of line


### Practice

A major consideration in any scatter plot is dealing with overplotting. 

You'll have to deal with overplotting when you have:

1. Large datasets.

2. Imprecise data and so points are not clearly separated on your plot.

3. Interval data (i.e. data appears at fixed values), or

4. Aligned data values on a single axis.

One very common technique that I'd recommend to always use when you have solid shapes it to use alpha blending (i.e. adding transparency). An alternative is to use hollow shapes. These are adjustments to make before even worrying about positioning. This addresses the first point as above, which you'll see again in the next exercise.

```{r}
# Basic scatter plot of wt on x-axis and mpg on y-axis; map cyl to col
ggplot(mtcars, aes(x = wt, y = mpg, color = cyl)) +
  geom_point(size = 4)

# Hollow circles - an improvement
ggplot(mtcars, aes(x = wt, y = mpg, color = cyl)) +
  geom_point(size = 4, shape = 1)

# Add transparency - very nice
ggplot(mtcars, aes(x = wt, y = mpg, color = cyl)) +
  geom_point(size = 4, alpha = 0.6)
```


```{r}
# Scatter plot: carat (x), price (y), clarity (color)
ggplot(diamonds, aes(x = carat, y = price, col = clarity)) + geom_point()

# Adjust for overplotting
ggplot(diamonds, aes(x = carat, y = price, col = clarity)) + geom_point(alpha = 0.5)

# Scatter plot: clarity (x), carat (y), price (color)
ggplot(diamonds, aes(x = clarity, y = carat, col = price)) + geom_point(alpha = 0.5)

# Dot plot with jittering
ggplot(diamonds, aes(x = clarity, y = carat, col = price)) + geom_point(alpha = 0.5,position = "jitter")

```

