---
title: "Datacamp_Data Visualization with ggplot2 (Part 1)__Data"
author: "dizhen"
date: "2019Äê4ÔÂ8ÈÕ"
output: html_document
---
# base package and ggplot2, part 1 - plot
These courses are about understanding data visualization in the context of the grammar of graphics. To gain a better appreciation of ggplot2 and to understand how it operates differently from base package, it's useful to make some comparisons.

In the video, you already saw one example of how to make a (poor) multivariate plot in base package. In this series of exercises you'll take a look at a better way using the equivalent version in ggplot2.

First, let's focus on base package. You want to make a plot of mpg (miles per gallon) against wt (weight in thousands of pounds) in the mtcars data frame, but this time you want the dots colored according to the number of cylinders, cyl. How would you do that in base package? You can use a little trick to color the dots by specifying a factor variable as a color. This works because factors are just a special class of the integer type.

Instructions

1. Using the base package plot(), make a scatter plot with mtcars$wt on the x-axis and mtcars$mpg on the y-axis, colored according to mtcars$cyl (use the col argument). You can specify data = but you'll just do it the long way here.

2. Add a new column, fcyl, to the mtcars data frame. This should be cyl converted to a factor.

3. Create a similar plot to instruction 1, but this time, use fcyl (which is cyl as a factor) to set the col.

```{r}
# Plot the correct variables of mtcars
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)

# Change cyl inside mtcars to a factor
mtcars$fcyl <- as.factor(mtcars$cyl)

# Make the same plot as in the first instruction
plot(mtcars$wt, mtcars$mpg, col = mtcars$fcyl)
```
# base package and ggplot2, part 2 - lm
If you want to add a linear model to your plot, shown right, you can define it with lm() and then plot the resulting linear model with abline(). However, if you want a model for each subgroup, according to cylinders, then you have a couple of options.

You can subset your data, and then calculate the lm() and plot each subset separately. Alternatively, you can vectorize over the cyl variable using lapply() and combine this all in one step. This option is already prepared for you.

The code to the right contains a call to the function lapply(), which you might not have seen before. This function takes as input a vector and a function. Then lapply() applies the function it was given to each element of the vector and returns the results in a list. In this case, lapply() takes each element of mtcars$cyl and calls the function defined in the second argument. This function takes a value of mtcars$cyl and then subsets the data so that only rows with cyl == x are used. Then it fits a linear model to the filtered dataset and uses that model to add a line to the plot with the abline() function.

Now that you have an interesting plot, there is a very important aspect missing - the legend!

In base package you have to take care of this using the legend() function. This has been done for you in the predefined code.

Instructions

1. Fill in the lm() function to calculate a linear model of mpg described by wt and save it as an object called carModel.

2. Draw the linear model on the scatterplot.

Write code that calls abline() with carModel as the first argument. Set the line type by passing the argument lty = 2.

Run the code that generates the basic plot and the call to abline() all at once by highlighting both parts of the script and hitting control/command + enter on your keyboard. These lines must all be run together in the DataCamp R console so that R will be able to find the plot you want to add a line to.

3. Run the code already given to generate the plot with a different model for each group. You don't need to modify any of this.

```{r}
# Use lm() to calculate a linear model and save it as carModel
carModel <- lm(mpg ~ wt, data = mtcars)

# Basic plot
mtcars$cyl <- as.factor(mtcars$cyl)
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)

# Call abline() with carModel as first argument and set lty to 2
abline(carModel, lty = 2)

# Plot each subset efficiently with lapply
# You don't have to edit this code
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)
lapply(mtcars$cyl, function(x) {
  abline(lm(mpg ~ wt, mtcars, subset = (cyl == x)), col = x)
  })

# This code will draw the legend of the plot
# You don't have to edit this code
legend(x = 5, y = 33, legend = levels(mtcars$cyl),
       col = 1:3, pch = 1, bty = "n")
```

# base package and ggplot2, part 3
In this exercise you'll recreate the base package plot in ggplot2.

The code for base R plotting is given at the top. The first line of code already converts the cyl variable of mtcars to a factor.


Instructions

1. Plot 1: add geom_point() in order to make a scatter plot.

2. Plot 2: copy and paste Plot 1.

3. Add a linear model for each subset according to cyl by adding a geom_smooth() layer.

4. Inside this geom_smooth(), set method to "lm" and se to FALSE.

5. Note: geom_smooth() will automatically draw a line per cyl subset. It recognizes the groups you want to identify by color in the aes() call within the ggplot() command.

6. Plot 3: copy and paste Plot 2.

7. Plot a linear model for the entire dataset, do this by adding another geom_smooth() layer.

8. Set the group aesthetic inside this geom_smooth() layer to 1. This has to be set within the aes() function.

9. Set method to "lm", se to FALSE and linetype to 2. These have to be set outside aes() of the geom_smooth().

10. Note: the group aesthetic will tell ggplot() to draw a single linear model through all the points.

```{r}
# Convert cyl to factor (don't need to change)
mtcars$cyl <- as.factor(mtcars$cyl)

# Example from base R (don't need to change)
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)
abline(lm(mpg ~ wt, data = mtcars), lty = 2)
lapply(mtcars$cyl, function(x) {
  abline(lm(mpg ~ wt, mtcars, subset = (cyl == x)), col = x)
  })
legend(x = 5, y = 33, legend = levels(mtcars$cyl),
       col = 1:3, pch = 1, bty = "n")

# Plot 1: add geom_point() to this command to create a scatter plot
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point()

# Plot 2: include the lines of the linear models, per cyl
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

# Plot 3: include a lm for the entire dataset in its whole
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE, linetype = 2)
```


# Variables to visuals, part 1
So far you've seen four different forms of the iris dataset: iris, iris.wide, iris.wide2 and iris.tidy. Don't let all these different forms confuse you! It's exactly the same data, just rearranged so that your plotting functions become easier.

To see this in action, consider the plot in the graphics device at right. Which form of the dataset would be the most appropriate to use here?

Instructions

1. Look at the structures of iris, iris.wide and iris.tidy using str().

2. Fill in the ggplot function with the appropriate data frame and variable names. The variable names of the aesthetics of the plot will match the ones you found using the str() command in the previous step.

```{r}
# Consider the structure of iris, iris.wide and iris.tidy (in that order)
str(iris)

# 'data.frame':	150 obs. of  5 variables:
# $ Species     : Factor w/ 3 levels "setosa","versicolor",..: 1 1 1 1 1 1 1 1 1 1 ...
# $ Sepal.Length: num  5.1 4.9 4.7 4.6 5 5.4 4.6 5 4.4 4.9 ...
# $ Sepal.Width : num  3.5 3 3.2 3.1 3.6 3.9 3.4 3.4 2.9 3.1 ...
# $ Petal.Length: num  1.4 1.4 1.3 1.5 1.4 1.7 1.4 1.5 1.4 1.5 ...
# $ Petal.Width : num  0.2 0.2 0.2 0.2 0.2 0.4 0.3 0.2 0.2 0.1 ...

str(iris.wide)

# 'data.frame':	300 obs. of  4 variables:
# $ Species: Factor w/ 3 levels "setosa","versicolor",..: 1 1 1 1 1 1 1 1 1 1 ...
# $ Part   : chr  "Petal" "Petal" "Petal" "Petal" ...
# $ Length : num  1.4 1.4 1.3 1.5 1.4 1.7 1.4 1.5 1.4 1.5 ...
# $ Width  : num  0.2 0.2 0.2 0.2 0.2 0.4 0.3 0.2 0.2 0.1 ...

str(iris.tidy)

# 'data.frame':	600 obs. of  4 variables:
# $ Species: Factor w/ 3 levels "setosa","versicolor",..: 1 1 1 1 1 1 1 1 1 1 ...
# $ Part   : chr  "Sepal" "Sepal" "Sepal" "Sepal" ...
# $ Measure: chr  "Length" "Length" "Length" "Length" ...
# $ Value  : num  5.1 4.9 4.7 4.6 5 5.4 4.6 5 4.4 4.9 ...


# Think about which dataset you would use to get the plot shown right
# Fill in the ___ to produce the plot given to the right
ggplot(iris.tidy, aes(x = Species, y = Value, col = Part)) +
  geom_jitter() +
  facet_grid(. ~ Measure)
```

# Variables to visuals, part 1b
In the last exercise you saw how iris.tidy was used to make a specific plot. It's important to know how to rearrange your data in this way so that your plotting functions become easier. In this exercise you'll use functions from the tidyr package to convert iris to iris.tidy.

The resulting iris.tidy data should look as follows:

      Species  Part Measure Value
    1  setosa Sepal  Length   5.1
    2  setosa Sepal  Length   4.9
    3  setosa Sepal  Length   4.7
    4  setosa Sepal  Length   4.6
    5  setosa Sepal  Length   5.0
    6  setosa Sepal  Length   5.4
    ...
You can have a look at the iris dataset by typing head(iris) in the console.

Note: If you're not familiar with %>%, gather() and separate(), you may want to take the Cleaning Data in R course. In a nutshell, a dataset is called tidy when every row is an observation and every column is a variable. The gather() function moves information from the columns to the rows. It takes multiple columns and gathers them into a single column by adding rows. The separate() function splits one column into two or more columns according to a pattern you define. Lastly, the %>% (or "pipe") operator passes the result of the left-hand side as the first argument of the function on the right-hand side.

Instructions

You'll use two functions from the tidyr package:

1. gather() rearranges the data frame by specifying the columns that are categorical variables with a - notation. Complete the command. Notice that only one variable is categorical in iris.

2. separate() splits up the new key column, which contains the former headers, according to .. The new column names "Part" and "Measure" are given in a character vector. Don't forget the quotes.

```{r}
# Load the tidyr package
library(tidyr)
head(iris.tidy)
head(iris)
# Fill in the ___ to produce to the correct iris.tidy dataset
iris.tidy <- iris %>%
  gather(key, Value, -Species) %>%
  separate(key, c("Part", "Measure"), "\\.")
```

# Variables to visuals, part 2
Here you'll take a look at another plot variant, shown on the right. Which of your data frames would be used to produce this plot?

Instructions

1. Look at the heads of iris, iris.wide and iris.tidy using head().

2. Fill in the ggplot function with the appropriate data frame and variable names. The names of the aesthetics of the plot will match with variable names in your dataset. The previous instruction will help you match variable names in datasets with the ones in the plot.

```{r}
# The 3 data frames (iris, iris.wide and iris.tidy) are available in your environment
# Execute head() on iris, iris.wide and iris.tidy (in that order)
head(iris)
head(iris.wide)
head(iris.tidy)

# Think about which dataset you would use to get the plot shown right
# Fill in the ___ to produce the plot given to the right
ggplot(iris.wide, aes(x = Length, y = Width, color = Part)) +
  geom_jitter() +
  facet_grid(. ~ Species)
```

# Variables to visuals, part 2b
In the last exercise you saw how iris.wide was used to make a specific plot. You also saw previously how you can derive iris.tidy from iris. Now you'll move on to produce iris.wide.

The head of the iris.wide should look like this in the end:

  Species  Part Length Width
1  setosa Petal    1.4   0.2
2  setosa Petal    1.4   0.2
3  setosa Petal    1.3   0.2
4  setosa Petal    1.5   0.2
5  setosa Petal    1.4   0.2
6  setosa Petal    1.7   0.4
...
You can have a look at the iris dataset by typing head(iris) in the console.

# Variables to visuals, part 2b
In the last exercise you saw how iris.wide was used to make a specific plot. You also saw previously how you can derive iris.tidy from iris. Now you'll move on to produce iris.wide.

The head of the iris.wide should look like this in the end:

  Species  Part Length Width
1  setosa Petal    1.4   0.2
2  setosa Petal    1.4   0.2
3  setosa Petal    1.3   0.2
4  setosa Petal    1.5   0.2
5  setosa Petal    1.4   0.2
6  setosa Petal    1.7   0.4
...
You can have a look at the iris dataset by typing head(iris) in the console.

Instructions

1. Before you begin, you need to add a new column called Flower that contains a unique identifier for each row in the data frame. This is because you'll rearrange the data frame afterwards and you need to keep track of which row, or which specific flower, each value came from. It's done for you, no need to add anything yourself.

2. gather() rearranges the data frame by specifying the columns that are categorical variables with a - notation. In this case, Species and Flower are categorical. Complete the command.

3. separate() splits up the new key column, which contains the former headers, according to .. The new column names "Part" and "Measure" are given in a character vector.

4. The last step is to use spread() to distribute the new Measure column and associated value column into two columns.

```{r}
# Load the tidyr package
library(tidyr)

# Add column with unique ids (don't need to change)
iris$Flower <- 1:nrow(iris)

# Fill in the ___ to produce to the correct iris.wide dataset
iris.wide <- iris %>%
  gather(key, value, -Species, -Flower) %>%
  separate(key, c("Part", "Measure"), "\\.") %>%
  spread(Measure, value)
```

