---
title: "Datacamp_Data Visualization with ggplot2 (Part 1)__Geometries"
author: "dizhen"
date: "2019/4/9"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scatter plot

1. Common plot types

* Scatter plots

points, jitter, abline

* Bar plots

histogram, bar, errorbar

* Line plots

line

2. Scatter plot

* Each geom has specific aesthetic mappings

* geom_point()

Essential: x, y; Optional: alpha, colour, fill, shape, size


3. Add layers

```{r}
library("ggplot2")
head(iris)
iris.summary <- aggregate(iris[1:4], list(iris$Species), mean) 
names(iris.summary)[1] <- "Species"
head(iris.summary)

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
  geom_point()+
  geom_point(data = iris.summary, shape = 15, size = 5)

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
  geom_point() +
  geom_point(data = iris.summary,
  shape = 21, size = 5, fill = "#00000080")
```

4. Cross hairs

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
  geom_point() +
  geom_vline(data = iris.summary,  aes(xintercept = Sepal.Length))

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
  geom_point() +
  geom_vline(data = iris.summary, aes(xintercept = Sepal.Length)) +
  geom_hline(data = iris.summary, aes(yintercept = Sepal.Width))

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
  geom_point() +
  geom_vline(data = iris.summary, linetype = 2,
  aes(xintercept = Sepal.Length, col = Species)) +
  geom_hline(data = iris.summary, linetype = 2,
  aes(yintercept = Sepal.Width, col = Species))
```

5. Remarks

* ggplot2 can also calculate statistics

* mean only = no good

6. jitter

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
  geom_point(position = "jitter")

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
  geom_jitter()

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
  geom_jitter(alpha = 0.6)

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
  geom_jitter(shape = 1)


```


### Practice

```{r}
# Shown in the viewer:
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_point()

# Solutions:
# 1 - With geom_jitter()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_jitter()

# 2 - Set width in geom_jitter()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_jitter(width = 0.1)

# 3 - Set position = position_jitter() in geom_point() ()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_point(position = position_jitter(0.1))
```


```{r}
library(readr)
Vocab <- read_csv("data/Vocab.csv")
head(Vocab)

# Examine the structure of Vocab
str(Vocab)

# Basic scatter plot of vocabulary (y) against education (x). Use geom_point()
ggplot(Vocab, aes(x = education, y = vocabulary)) + geom_point()

# Use geom_jitter() instead of geom_point()
ggplot(Vocab, aes(x = education, y = vocabulary)) + geom_jitter()

# Using the above plotting command, set alpha to a very low 0.2
ggplot(Vocab, aes(x = education, y = vocabulary)) + geom_jitter(alpha = 0.2)

# Using the above plotting command, set the shape to 1
ggplot(Vocab, aes(x = education, y = vocabulary)) + geom_jitter(alpha = 0.2, shape = 1)

```


## Bar plots

1. Common plot types

* Scatter plots

points, jitter, abline

* Bar plots

histogram, bar, errorbar

* Line plots

line

2. Histogram

Features:

* Many ways to do binning

* No space between bars

* x axis labels between the bars

* Summary table in the background


```{r}
ggplot(iris, aes(x = Sepal.Width)) +
  geom_histogram()
diff(range(iris$Sepal.Width)) / 30

ggplot(iris, aes(x = Sepal.Width)) +
  geom_histogram(binwidth = 0.1)

# internal data frame with y re-labeled
ggplot(iris, aes(x = Sepal.Width)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.1)

# different species
ggplot(iris, aes(x = Sepal.Width, fill = Species)) +
  geom_histogram(binwidth = 0.1)

ggplot(iris, aes(x = Sepal.Width, fill = Species)) +
  geom_histogram(binwidth = 0.1, position = "stack")

ggplot(iris, aes(x = Sepal.Width, fill = Species)) +
  geom_histogram(binwidth = 0.1, position = "dodge")

ggplot(iris, aes(x = Sepal.Width, fill = Species)) +
  geom_histogram(binwidth = 0.1, position = "fill")
```

3. Bar plot

* geom_bar()

* All positions from before available

* Two types

Absolute counts and Distributions

```{r}
library(plyr)
library(reshape2)
iris_melted <- melt(iris, value.name = "Value", variable.name = "Measure") 
iris_summ <- ddply(iris_melted[iris_melted$Measure == "Sepal.Width",], "Species", summarise, avg = mean(Value), stdev = sd(Value))
str(iris_summ)


ggplot(iris_summ, aes(x = Species, y = avg)) +
  geom_bar(stat = "identity")

ggplot(iris_summ, aes(x = Species, y = avg)) +
  geom_bar(stat = "identity", fill = "grey50") +
  geom_errorbar(aes(ymin = avg - stdev, ymax = avg + stdev),width = 0.2)
```


### Practice

**The x axis/aesthetic**: The documentation for *geom_histogram()* states the argument *stat = "bin"* as a default. Recall that histograms cut up a continuous variable into discrete bins - that's what the stat "bin" is doing. You always get 30 evenly-sized bins by default, which is specified with the default argument *binwidth = range/30*. This is a pretty good starting point if you don't know anything about the variable being ploted and want to start exploring.

**The y axis/aesthetic**: *geom_histogram()* only requires one aesthetic: *x*. But there is clearly a *y* axis on your plot, so where does it come from? Actually, there is a variable mapped to the y aesthetic, it's called *..count..*. When *geom_histogram()* executed the binning statistic (see above), it not only cut up the data into discrete bins, but it also counted how many values are in each bin. So there is an internal data frame where this information is stored. The *..* calls the variable *count* from this internal data frame. This is what appears on the *y* aesthetic. But it gets better! The density has also been calculated. This is the proportional frequency of this bin in relation to the whole data set. You use *..density..* to access this information.

```{r}
# 1 - Make a univariate histogram
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram()

# 2 - Plot 1, plus set binwidth to 1 in the geom layer
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth = 1)

# 3 - Plot 2, plus MAP ..density.. to the y aesthetic (i.e. in a second aes() function)
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth = 1, aes(y = ..density..))

# 4 - plot 3, plus SET the fill attribute to "#377EB8"
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth = 1, aes(y = ..density..),fill = "#377EB8")


```

Three position arguments will be introduced here:

* stack: place the bars on top of each other. Counts are used. This is the default position.

* fill: place the bars on top of each other, but this time use proportions.

* dodge: place the bars next to each other. Counts are used.

```{r}
# Draw a bar plot of cyl, filled according to am
ggplot(mtcars, aes(x = cyl, fill = as.factor(am))) +
  geom_bar()

# Change the position argument to stack
ggplot(mtcars, aes(x = cyl, fill = as.factor(am))) +
  geom_bar(position = "stack")

# Change the position argument to fill
ggplot(mtcars, aes(x = cyl, fill = as.factor(am))) +
  geom_bar(position = "fill")

# Change the position argument to dodge
ggplot(mtcars, aes(x = cyl, fill = as.factor(am))) +
  geom_bar(position = "dodge")

```


Instead of using *position = "dodge"* you're going to use *position_dodge()*, like you did with *position_jitter()* in the Scatter plots and jittering exercise. Here, you'll save this as an object, *posn_d*, so that you can easily reuse it.

Remember, the reason you want to use *position_dodge()* (and *position_jitter()*) is to specify how much dodging (or jittering) you want.

```{r}
# 1 - The last plot form the previous exercise
ggplot(mtcars, aes(x = cyl, fill = as.factor(am))) +
  geom_bar(position = "dodge")

# 2 - Define posn_d with position_dodge()
posn_d <- position_dodge(width=0.2)

# 3 - Change the position argument to posn_d
ggplot(mtcars, aes(x = cyl, fill = as.factor(am))) +
  geom_bar(position = posn_d)

# 4 - Use posn_d as position and adjust alpha to 0.6
ggplot(mtcars, aes(x = cyl, fill = as.factor(am))) +
  geom_bar(position = posn_d,alpha = 0.6)


```

Using the same data and base layers as in the previous two plots, create a plot with a *geom_freqpoly()*. Because you're no longer working with bars, change the *aes()* function: *cyl* should be mapped onto *color*, not onto *fill*.


```{r}
# A basic histogram, add coloring defined by cyl
ggplot(mtcars, aes(mpg,fill = as.factor(cyl))) +
  geom_histogram(binwidth = 1)

# Change position to identity
ggplot(mtcars, aes(mpg,fill = as.factor(cyl))) +
  geom_histogram(binwidth = 1,position = "identity")

# Change geom to freqpoly (position is identity by default)
ggplot(mtcars, aes(mpg,col = as.factor(cyl))) +
  geom_freqpoly(binwidth = 1,position = "identity")

```


You can set the color palette used to fill the bars with *scale_fill_brewer()*. For a full list of possible color sets, have a look at *?brewer.pal*.

```{r}
# Example of how to use a brewed color palette
ggplot(mtcars, aes(x = cyl, fill = as.factor(am))) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1")

# Use str() on Vocab to check out the structure
str(Vocab)

# Plot education on x and vocabulary on fill
# Use the default brewed color palette
ggplot(Vocab, aes(x = education, fill = as.factor(vocabulary))) +
  geom_bar(position = "fill") +
  scale_fill_brewer(palette = "Set1")


```

For continuous data, the default *RColorBrewer* palette that *scale_fill_brewer()* calls is "*Blues*". There are only 9 colours in the palette, and since you have 11 categories, your plot looked strange.

Here you'll manually create a color palette that can generate all the colours you need. To do this you'll use a function called *colorRampPalette()*.

The input is a character vector of 2 or more colour values, e.g. "*#FFFFFF*" (white) and "*#0000FF*" (pure blue). (See this exercise for a discussion on hexadecimal codes).

The output is itself a function! So when you assign it to an object, that object should be used as a function.

```{r}
new_col <- colorRampPalette(c("#FFFFFF", "#0000FF"))
new_col(4) # the newly extrapolated colours
munsell::plot_hex(new_col(4)) # Quick and dirty plot
```

*new_col()* is a function that takes one argument: the number of colours you want to extrapolate. You want to use nicer colours, so we've assigned the entire "*Blues*" colour palette from the *RColorBrewer* package to the character vector *blues.*



```{r}
# Final plot of last exercise
ggplot(Vocab, aes(x = education, fill = as.factor(vocabulary))) +
  geom_bar(position = "fill") +
  scale_fill_brewer()

# Definition of a set of blue colors
library("RColorBrewer")
blues <- brewer.pal(9, "Blues") # from the RColorBrewer package

# 1 - Make a color range using colorRampPalette() and the set of blues
blue_range <- colorRampPalette(blues)

# 2 - Use blue_range to adjust the color of the bars, use scale_fill_manual()
ggplot(Vocab, aes(x = education, fill = as.factor(vocabulary))) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = blue_range(11))

```



```{r}
# 1 - Basic histogram plot command
ggplot(mtcars, aes(mpg)) +
  geom_histogram(binwidth = 1)

# 2 - Plot 1, Expand aesthetics: am onto fill
ggplot(mtcars, aes(mpg,fill = as.factor(am))) +
  geom_histogram(binwidth = 1)

# 3 - Plot 2, change position = "dodge"
ggplot(mtcars, aes(mpg,fill = as.factor(am))) +
  geom_histogram(binwidth = 1,position = "dodge")

# 4 - Plot 3, change position = "fill"
ggplot(mtcars, aes(mpg,fill = as.factor(am))) +
  geom_histogram(binwidth = 1,position = "fill")

# 5 - Plot 4, plus change position = "identity" and alpha = 0.4
ggplot(mtcars, aes(mpg,fill = as.factor(am))) +
  geom_histogram(binwidth = 1,position = "identity",alpha = 0.4)

# 6 - Plot 5, plus change mapping: cyl onto fill
ggplot(mtcars, aes(mpg,fill = as.factor(cyl))) +
  geom_histogram(binwidth = 1,position = "identity",alpha = 0.4)

```
## Line plots

```{r eval = FALSE}

ggplot(beaver, aes(x = time, y = temp)) +
  geom_line()

ggplot(beaver, aes(x = time, y = temp, col = factor(active))) +
  geom_line()

# Aesthetics - categorical variables

# linetype
ggplot(fish, aes(x = Year, y = Capture, linetype = Species)) +
  geom_line()

# size
ggplot(fish, aes(x = Year, y = Capture, size = Species)) +
  geom_line()

# color
ggplot(fish, aes(x = Year, y = Capture, color = Species)) +
  geom_line()

# fill
ggplot(fish, aes(x = Year, y = Capture, fill = Species)) +
  geom_area()
ggplot(fish, aes(x = Year, y = Capture, fill = Species)) +
  geom_area(position = "fill")

# geom_ribbon
ggplot(fish, aes(x = Year, y = Capture, fill = Species)) +
  geom_ribbon(aes(ymax = Capture, ymin = 0), alpha = 0.3)

```




### Practice



```{r}
  # Print out head of economics
head(economics)

# Plot unemploy as a function of date using a line plot
ggplot(economics, aes(x = date, y = unemploy)) + geom_line()

# Adjust plot to represent the fraction of total population that is unemployed
ggplot(economics, aes(x = date, y = unemploy/pop)) + geom_line()

```

The *geom_rect()* command shouldn't inherit aesthetics from the base *ggplot()* command it belongs to. It would result in an error, since you're using a different dataset and it doesn't contain *unemploy* or *pop.* That's why you should specify *inherit.aes = FALSE* in *geom_rect()*

```{r}
# Basic line plot
ggplot(economics, aes(x = date, y = unemploy/pop)) +
  geom_line()

# Expand the following command with geom_rect() to draw the recess periods
begin <- c("1969-12-01","1973-11-01","1980-01-01","1981-07-01","1990-07-01","2001-03-01")
end <- c("1970-11-01","1975-03-01","1980-07-01","1982-11-01","1991-03-01","2001-11-01")
recess <- data.frame(begin, end)
recess$begin <- as.Date(recess$begin,format = "%Y-%m-%d")
recess$end <- as.Date(recess$end,format = "%Y-%m-%d")
# mutate(recess, begin = as.Date(begin,format = "%Y-%m-%d"))
str(recess)

ggplot(economics, aes(x = date, y = unemploy/pop)) +
  geom_rect(data = recess,
         aes(xmin = begin, xmax = end, ymin = -Inf, ymax = +Inf),
         inherit.aes = FALSE, fill = "red", alpha = 0.2) +
  geom_line()
```


The dataset you'll use contains the global capture rates of seven salmon species from 1950 - 2010.

In your workspace, the following dataset is available:

fish.species: Each variable (column) is a Salmon Species and each observation (row) is one Year.
To get a multiple time series plot, however, both Year and Species should be in their own column. You need tidy data: one variable per column. Once you have that you can get the plot shown in the viewer by mapping Year to the x aesthetic and Species to the color aesthetic.

You'll use the gather() function of the tidyr package, which is already loaded for you.

1. Use gather() to move from fish.species to a tidy data frame, fish.tidy. This data frame should have three columns: Year (int), Species (factor) and Capture (int).

2. gather() takes four arguments: the original data frame (fish.species), the name of the key column (Species), the name of the value column (Capture) and the name of the grouping variable, with a minus in front (-Year). They can all be specified as object names.



```{r}
fish.species <- read_csv("data/fish.species.csv")
fish.species <- fish.species[c("Year","Pink","Chum","Sockeye","Coho","Rainbow","Chinook","Atlantic")]
head(fish.species)
# Check the structure as a starting point
str(fish.species)

# Use gather to go from fish.species to fish.tidy
library("tidyr")
library("reshape2")
fish.tidy <- gather(fish.species, Species, Capture, -Year)
head(fish.tidy)
```


```{r}
# Recreate the plot shown on the right
ggplot(fish.tidy, aes(x = Year, y = Capture,col = Species)) + geom_line()
```


