---
title: "Datacamp_Data Visualization with ggplot2 (Part 1)_Introduction"
author: "dizhen"
date: "2019/4/6"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

```{r}
# Load the ggplot2 package
library(ggplot2)

# Explore the mtcars data frame with str()
str(mtcars)
head(mtcars)
```

```{r}
ggplot(mtcars, aes(x = cyl, y = mpg)) +
  geom_point()

# cyl is treated as factor
ggplot(mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_point()
```

## Grammer of Graphics

Essential Grammatic Elements:

1. Data: The dataset being plotted. (variables of interest)

2. Asethetics: The scales onto which we map our data. (x-axis, y-axis; colour, fill; size, labels; alpha, shape; line width, line type)

3. Geometries: The visual element used for out data. (point, line, histogram, bar, boxplot)

4. Facets: plotting small multiples (columns, rows)

5. Statistics: Representations of our data to aid understanding (binning, smoothing, descriptive, inferential)

6. Coordinates: The space on which the data will be plotted (cartesian, fixed, polar, limits)

7. Themes: All non-data ink. (non-data ink)

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()

# color changes by disp (continuous)
ggplot(mtcars, aes(x = wt, y = mpg, color = disp)) +
  geom_point()

# size changes by disp (continuous)
ggplot(mtcars, aes(x = wt, y = mpg, size = disp)) +
  geom_point()
```

As "shape" only makes sense with categorical data, and *disp* is cotinuous. So the code below is incorrect.
```{r eval = FALSE}
ggplot(mtcars, aes(x = wt, y = mpg, shape = disp)) +
  geom_point()
```

## ggplot 2

```{r}
# Explore the diamonds data frame with str()
str(diamonds)
head(diamonds)
```

```{r}
# Add geom_point() for scatter plot
ggplot(diamonds, aes(x = carat, y = price)) + geom_point()

# Add geom_point() and geom_smooth() 
ggplot(diamonds, aes(x = carat, y = price))+ geom_point() + geom_smooth()

```


```{r}
# 1 - The plot you created in the previous exercise
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth()

# 2 - show only the smooth line
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_smooth()


# 3 - assign the correct value to col in aes()
# different color of smooths by different clarity
ggplot(diamonds, aes(x = carat, y = price, color = clarity)) +
  geom_smooth()


# 4 -  Plot only the points with argument alpha.
# different color of points by different clarity
# change the transparent as 40% visible
ggplot(diamonds, aes(x = carat, y = price, color = clarity)) +
  geom_point(alpha = 0.4) 
```

```{r}
# Create the object containing the data and aes layers: dia_plot
dia_plot <- ggplot(diamonds, aes(x = carat, y = price))

# Add a geom layer with + and geom_point()
dia_plot + geom_point()

# Add the same geom layer, but with aes() inside
# can also call aes() within the geom_point() function
dia_plot + geom_point(aes(color = clarity))

```

```{r}
# 1 - The dia_plot object has been created 
dia_plot <- ggplot(diamonds, aes(x = carat, y = price))

# 2 - Expand dia_plot by adding geom_point() with alpha set to 0.2
dia_plot <- dia_plot + geom_point(alpha = 0.2)

# 3 - Plot dia_plot with additional geom_smooth() with se set to FALSE
# do not include standard error
dia_plot + geom_smooth(se = FALSE)

# 4 - Copy the command from above and add aes() with the correct mapping to geom_smooth()
dia_plot + geom_smooth(aes(col = clarity), se = FALSE)
```


