---
title: "Datacamp_Data Visualization with ggplot2 (Part 1)__qplot and wrap-up"
author: "dizhen"
date: '2020/4/9'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## qplot

```{r}
library("ggplot2")

# ggplot
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) +
 geom_point()

# qplot
qplot(Sepal.Length, Sepal.Width, data = iris) 

# base
plot(iris$Sepal.Length, iris$Sepal.Width) 
plot(Sepal.Width ~ Sepal.Length, data = iris)
```


```{r}
# shape = Species
qplot(Sepal.Length, Sepal.Width, data = iris, shape = Species) 

# size = Species
qplot(Sepal.Length, Sepal.Width, data = iris, size = Species) 

# col = Species
qplot(Sepal.Length, Sepal.Width, data = iris, col = Species) 

# geom argument
qplot(Sepal.Length, Sepal.Width, data = iris, col = Species,
  geom = "jitter")

# position argument
qplot(Sepal.Length, Sepal.Width, data = iris, col = Species,
  position = "jitter") 

# alpha
qplot(Sepal.Length, Sepal.Width, data = iris, col = Species,
  position = "jitter", alpha = 0.5) 
qplot(Sepal.Length, Sepal.Width, data = iris, col = Species,
  position = "jitter", alpha = I(0.5))

 
 
```

```{r}
library(tidyr)
library(tidyverse)
iris.tidy <- iris %>%
  gather(key, Value, -Species) %>%
  separate(key, c("Part", "Measure"), "\\.")
str(iris)

# continuous vs categorical
qplot(Species, Value, data = iris.tidy, col = Part) 

# position = "jitter"
qplot(Species, Value, data = iris.tidy, col = Part,
  position = "jitter")

# jitter manually
posn.j <- position_jitter(0.1) 
qplot(Species, Value, data = iris.tidy, col = Part,
  position = posn.j) 

# comparison
posn.j <- position_jitter(0.1)
qplot(Species, Value, data = iris.tidy, col = Part,
  position = posn.j) # Fine for easy plot
ggplot(iris.tidy, aes(x = Species, y = Value, col = Part)) +
  geom_point(position = posn.j) # very flexible

```

### Practice


```{r}
# The old way (shown)
plot(mpg ~ wt, data = mtcars) # formula notation
with(mtcars, plot(wt, mpg)) # x, y notation

# Using ggplot:
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()

# Using qplot:
qplot(wt, mpg, data = mtcars)

```



```{r}
# basic qplot scatter plot:
qplot(wt, mpg, data = mtcars)

# Categorical variable mapped onto size:
# cyl
qplot(wt, mpg, data = mtcars, size = factor(cyl))

# gear
qplot(wt, mpg, data = mtcars, size = factor(gear))

# Continuous variable mapped onto col:
# hp
qplot(wt, mpg, data = mtcars, col = hp)

# qsec
qplot(wt, mpg, data = mtcars, col = qsec)
```



```{r}
# qplot() with x only
qplot(x = factor(cyl), data = mtcars)

# qplot() with x and y
qplot(x = factor(cyl), y = factor(vs),data = mtcars)

# qplot() with geom set to jitter manually
qplot(x = factor(cyl), y = factor(vs),data = mtcars,geom ="jitter")

```

Some naming conventions:

* Scatter plots:

Continuous x, continuous y.

* Dot plots:

Categorical x, continuous y.

You use geom_point() for both plot types. Jittering position is set in the geom_point() layer.

However, to make a "true" dot plot, you can use geom_dotplot(). The difference is that unlike geom_point(), geom_dotplot() uses a binning statistic. Binning means to cut up a continuous variable (the y in this case) into discrete "bins". You already saw binning with geom_histogram() (see this exercise for a refresher).

One thing to notice is that geom_dotplot() uses a different plotting symbol to geom_point(). For these symbols, the color aesthetic changes the color of its border, and the fill aesthetic changes the color of its interior.


```{r}
# cyl and am are factors, wt is numeric
class(mtcars$cyl)
class(mtcars$am)
class(mtcars$wt)

# "Basic" dot plot, with geom_point():
ggplot(mtcars, aes(cyl, wt, col = am)) +
  geom_point(position = position_jitter(0.2, 0))

# 1 - "True" dot plot, with geom_dotplot():
ggplot(mtcars, aes(cyl, wt, fill = am)) +
  geom_dotplot(binaxis = "y", stackdir = "center")

# 2 - qplot with geom "dotplot", binaxis = "y" and stackdir = "center"
qplot(
  x = cyl, y = wt,
  data = mtcars,
  fill = am,
  geom = "dotplot",
  binaxis = "y",
  stackdir = "center"
)
```
**Chicken weight**

The ChickWeight dataset is a data frame which represents the progression of weight of several chicks. The little chicklings are each given a specific diet. There are four types of diet and the farmer wants to know which one fattens the chicks the fastest.


```{r}
# ChickWeight is available in your workspace
# 1 - Check out the head of ChickWeight
head(ChickWeight)

# 2 - Basic line plot
ggplot(ChickWeight, aes(x = Time, y = weight)) +
  geom_line(aes(group = Chick))

# 3 - Take plot 2, map Diet onto col.
ggplot(ChickWeight, aes(x = Time, y = weight, col = Diet)) +
  geom_line(aes(group = Chick))


# 4 - Take plot 3, add geom_smooth()
ggplot(ChickWeight, aes(x = Time, y = weight, col = Diet)) +
  geom_line(aes(group = Chick), alpha = 0.3) + geom_smooth(lwd = 2, se = FALSE)


```
**Titanic**

You've watched the movie Titanic by James Cameron (1997) again and after a good portion of sobbing you decide to investigate whether you'd have a chance of surviving this disaster.

```{r}
library(readr)
titanic <- read_csv("data/titanic.csv")
titanic <- titanic[,c("Survived","Pclass","Sex","Age")]
titanic <- na.omit(titanic)
titanic$Survived <- as.integer(titanic$Survived)
titanic$Pclass <- as.integer(titanic$Pclass)
titanic$Sex <- as.factor(titanic$Sex)


# 1 - Check the structure of titanic
str(titanic)
```

```{r}

# 2 - Use ggplot() for the first instruction
ggplot(titanic, aes(x = Pclass, fill = Sex)) +
  geom_bar(position = "dodge")

# 3 - Plot 2, add facet_grid() layer
ggplot(titanic, aes(x = Pclass, fill = Sex)) +
  geom_bar(position = "dodge") + facet_grid(.~Survived)

# 4 - Define an object for position jitterdodge, to use below
posn.jd <- position_jitterdodge(0.5, 0, 0.6)

# 5 - Plot 3, but use the position object from instruction 4
ggplot(titanic, aes(x = Pclass, y = Age, col = Sex)) +
  geom_point(position = posn.jd, size = 3, alpha = 0.5) + facet_grid(.~Survived)
```

