---
title: 'Datacamp_Supervised Learning in R: Classification__Chapter 1: k-Nearest Neighbors
  (kNN)'
author: "dizhen"
date: "2019Äê4ÔÂ24ÈÕ"
output: html_document
---
# Recognizing a road sign with kNN
After several trips with a human behind the wheel, it is time for the self-driving car to attempt the test course alone.

As it begins to drive away, its camera captures the following image:

Stop Sign

Can you apply a kNN classifier to help the car recognize this sign?

The dataset signs is loaded in your workspace along with the dataframe next_sign, which holds the observation you want to classify.

Instructions

Load the class package.
Create a vector of sign labels to use with kNN by extracting the column sign_type from signs.
Identify the next_sign using the knn() function.
Set the train argument equal to the signs data frame without the first column.
Set the test argument equal to the data frame next_sign.
Use the vector of labels you created as the cl argument.

```{r}
# Load the 'class' package
library(class)

# Create a vector of labels
sign_types <- signs$sign_type

# Classify the next sign observed
knn(train = signs[-1], test = next_sign, cl = sign_types)
```

# Exploring the traffic sign dataset
To better understand how the knn() function was able to classify the stop sign, it may help to examine the training dataset it used.

Each previously observed street sign was divided into a 4x4 grid, and the red, green, and blue level for each of the 16 center pixels is recorded as illustrated here.

The result is a dataset that records the sign_type as well as 16 x 3 = 48 color properties of each sign.

Instructions

Use the str() function to examine the signs dataset.
Use table() to count the number of observations of each sign type by passing it the column containing the labels.
Run the provided aggregate() command to see whether the average red level might vary by sign type.

```{r}
# Examine the structure of the signs dataset
str(signs)

# Count the number of signs of each type
table(signs$sign_type)

# Check r10's average red level by sign type
aggregate(r10 ~ sign_type, data = signs, mean)
```

# Classifying a collection of road signs
Now that the autonomous vehicle has successfully stopped on its own, your team feels confident allowing the car to continue the test course.

The test course includes 58 additional road signs divided into three types:

At the conclusion of the trial, you are asked to measure the car's overall performance at recognizing these signs.

The class package and the dataset signs are already loaded in your workspace. So is the dataframe test_signs, which holds a set of observations you'll test your model on.

Instructions

Classify the test_signs data using knn().
Set train equal to the observations in signs without labels.
Use test_signs for the test argument, again without labels.
For the cl argument, use the vector of labels provided for you.
Use table() to explore the classifier's performance at identifying the three sign types (the confusion matrix).
Create the vector signs_actual by extracting the labels from test_signs.
Pass the vector of predictions and the vector of actual signs to table() to cross tabulate them.
Compute the overall accuracy of the kNN learner using the mean() function.

```{r}
# Use kNN to identify the test road signs
sign_types <- signs$sign_type
signs_pred <- knn(train = signs[-1], test = test_signs[-1], cl = sign_types)

# Create a confusion matrix of the predicted versus actual values
signs_actual <- test_signs$sign_type
table(signs_pred, signs_actual)

# Compute the accuracy
mean(signs_pred == signs_actual)
```

# Testing other 'k' values
By default, the knn() function in the class package uses only the single nearest neighbor.

Setting a k parameter allows the algorithm to consider additional nearby neighbors. This enlarges the collection of neighbors which will vote on the predicted class.

Compare k values of 1, 7, and 15 to examine the impact on traffic sign classification accuracy.

The class package is already loaded in your workspace along with the datasets signs, signs_test, and sign_types. The object signs_actual holds the true values of the signs.

Instructions

Compute the accuracy of the default k = 1 model using the given code, then find the accuracy of the model using mean() to compare signs_actual and the model's predictions.
Modify the knn() function call by setting k = 7 and again find accuracy value.
Revise the code once more by setting k = 15, plus find the accuracy value one more time.

```{r}
# Compute the accuracy of the baseline model (default k = 1)
k_1 <- knn(train = signs[-1], test = signs_test[-1], cl = sign_types)
mean(signs_actual == k_1)

# Modify the above to set k = 7
k_7 <- knn(train = signs[-1], test = signs_test[-1], cl = sign_types, k = 7)
mean(signs_actual == k_7)

# Set k = 15 and compare to the above
k_15 <- knn(train = signs[-1], test = signs_test[-1], cl = sign_types, k = 15)
mean(signs_actual == k_15)
```
# Seeing how the neighbors voted
When multiple nearest neighbors hold a vote, it can sometimes be useful to examine whether the voters were unanimous or widely separated.

For example, knowing more about the voters' confidence in the classification could allow an autonomous vehicle to use caution in the case there is any chance at all that a stop sign is ahead.

In this exercise, you will learn how to obtain the voting results from the knn() function.

The class package has already been loaded in your workspace along with the datasets signs, sign_types. and signs_actual.

Instructions

Build a kNN model with the prob = TRUE parameter to compute the vote proportions. Set k = 7.
Use the attr() function to obtain the vote proportions for the predicted class. These are stored in the attribute "prob".
Examine the first several vote outcomes and percentages using the head() function to see how the confidence varies from sign to sign.

```{r}
# Use the prob parameter to get the proportion of votes for the winning class
sign_pred <- knn(train = signs[-1], test = signs_test[-1], cl = sign_types, k = 7, prob = TRUE)

# Get the "prob" attribute from the predicted classes
sign_prob <- attr(sign_pred, "prob")

# Examine the first several predictions
head(sign_pred)

# Examine the proportion of votes for the winning class
head(sign_prob)
```

