---
title: "Datacamp_Data Visualization with ggplot2 (Part 2)__Statistics"
author: "dizhen"
date: "2019/4/10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Statistics with Geoms

1. Statistics Layer

Two categories of functions 

* Called from within a geom

* Called independently 

2. geom_ <-> stat_

* geom_histogram() <-> stat_bin() 

* geom_bar() <-> stat_bin()

* geom_freqpoly() <-> stat_bin()

* geom_smooth() <-> stat_smooth()

3. Other stat_ functions

stat_ <-> geom_

stat_bin() <-> geom_histogram()

stat_bin() <-> geom_bar()

stat_bin() <-> geom_freqpoly()

stat_smooth() <-> geom_smooth()

stat_boxplot() <-> geom_boxplot()

stat_bindot() <-> geom_dotplot()

stat_bin2d() <-> geom_bin2d()

stat_binhex() <-> geom_hex()

stat_contour() <-> geom_contour()

stat_quantile() <-> geom_quantile()

stat_sum() <-> geom_count()



```{r}

library("ggplot2")

# geom_ <-> stat_
p <- ggplot(iris, aes(x = Sepal.Width))
p + geom_histogram()
p + geom_bar()
p + stat_bin()

# grouping by fill
ggplot(mtcars, aes(x = factor(cyl), fill = factor(am))) +
  geom_bar()
# ggplot(mtcars, aes(x = factor(cyl), fill = factor(am))) +
#   stat_bin() # stat_bin require a continuous x variable

ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(fill = "skyblue") +
  geom_freqpoly(col = "red") 

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
  geom_point() +
  geom_smooth(method = "lm",se = FALSE, span = 0.4, fullrange = TRUE)


```

### Practice

```{r}

# Explore the mtcars data frame with str()
str(mtcars)

# A scatter plot with LOESS smooth
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + geom_smooth(method = "loess")

# A scatter plot with an ordinary Least Squares linear model
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + geom_smooth(method = "lm") 

# The previous plot, without CI ribbon
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + geom_smooth(method = "lm",se = FALSE)

# The previous plot, without points
ggplot(mtcars, aes(x = wt, y = mpg)) + geom_smooth(method = "lm",se = FALSE)
```

For 2, add a group aesthetic inside the aes() of this new stat_smooth(), set it to a dummy variable, 1.

```{r}

# 1 - Define cyl as a factor variable
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)

# 2 - Plot 1, plus another stat_smooth() containing a nested aes()
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE) + stat_smooth(method = "lm", se = FALSE, aes(group = 1))

```


Plot 1: Recall that LOESS smoothing is a non-parametric form of regression that uses a weighted, sliding-window, average to calculate a line of best fit. We can control the size of this window with the span argument.

Plot 2: In this plot, we set a linear model for the entire dataset as well as each subgroup, defined by cyl. In the second stat_smooth(),

Set method to "loess" (this is the default with a small (n < 1000) data set, but we will specify it explicitly).
Add span, set it to 0.7.

Plot 3: Plot 2 presents a problem because there is a black line on our plot that is not included in the legend. To get this, we need to map something to col as an aesthetic, not just set col as an attribute.

Add col to the aes() function in the second stat_smooth(), set it to "All". This will name the line properly.

Remove the col attribute in the second stat_smooth(). Otherwise, it will overwrite the col aesthetic.

Plot 4: Now we should see our "All" model in the legend, but it's not black anymore.

Add a scale layer: scale_color_manual() with the first argument set to "Cylinders" and values set to the predefined myColors variable.

```{r}
# Plot 1: change the LOESS span
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  # Add span below
  geom_smooth(se = FALSE, span = 0.7)

# Plot 2: Set the second stat_smooth() to use LOESS with a span of 0.7
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE) +
  # Change method and add span below
  stat_smooth(method = "loess", aes(group = 1),
              se = FALSE, col = "black", span = 0.7)

# Plot 3: Set col to "All", inside the aes layer of stat_smooth()
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE) +
  stat_smooth(method = "loess",
              # Add col inside aes()
              aes(group = 1, col = "All"),
              # Remove the col argument below
              se = FALSE, span = 0.7)

# Plot 4: Add scale_color_manual to change the colors
library("RColorBrewer")
myColors <- c(brewer.pal(3, "Dark2"), "black")
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, span = 0.7) +
  stat_smooth(method = "loess", 
              aes(group = 1, col="All"), 
              se = FALSE, span = 0.7) +
  # Add correct arguments to scale_color_manual
  scale_color_manual("Cylinders", values = myColors)
```


```{r}
library(carData)
str(Vocab)

# Plot 1: Jittered scatter plot, add a linear model (lm) smooth
ggplot(Vocab, aes(x = education, y = vocabulary)) +
  geom_jitter(alpha = 0.2) +
  stat_smooth(method = "lm",se = FALSE) # smooth

# Plot 2: points, colored by year
ggplot(Vocab, aes(x = education, y = vocabulary, col = year)) +
  geom_jitter(alpha = 0.2) 

# Plot 3: lm, colored by year
ggplot(Vocab, aes(x = education, y = vocabulary, col = factor(year))) +
  stat_smooth(method = "lm",se = FALSE) # smooth
  
# Plot 4: Set a color brewer palette
ggplot(Vocab, aes(x = education, y = vocabulary, col = factor(year))) +
  stat_smooth(method = "lm",se = FALSE) +  # smooth
  scale_color_brewer()  # colors

# Plot 5: Add the group aes, specify alpha and size
ggplot(Vocab, aes(x = education, y = vocabulary, col = year, group = factor(year))) +
  stat_smooth(method = "lm", se = FALSE, alpha = 0.6, size = 2) +
  scale_color_gradientn(colors = brewer.pal(9, "YlOrRd"))
```

Use stat_quantile() to apply a quantile regression (method rq). By default, the 1st, 2nd (i.e. median), and 3rd quartiles are modeled as a response to the predictor variable.Specific quantiles can be specified with the quantiles argument.

The first plot may be a mess, because there are three quartiles drawn by default.

```{r}
# Use stat_quantile instead of stat_smooth
ggplot(Vocab, aes(x = education, y = vocabulary, col = year, group = factor(year))) +
  stat_quantile(alpha = 0.6, size = 2) +
  scale_color_gradientn(colors = brewer.pal(9,"YlOrRd"))

# Set quantile to 0.5
ggplot(Vocab, aes(x = education, y = vocabulary, col = year, group = factor(year))) +
  stat_quantile(alpha = 0.6, size = 2,quantiles = 0.5) +
  scale_color_gradientn(colors = brewer.pal(9,"YlOrRd"))
```


Another useful stat function is stat_sum(). This function calculates the total number of overlapping observations and is another good alternative to overplotting.

1. Add stat_sum() to this plotting object p. This maps the overall count of each dot onto size. 

2. Add the size scale with the generic scale_size() function. Use range to set the minimum and maximum dot sizes as c(1,10).

```{r}
# Plot 1: Jittering only
p <- ggplot(Vocab, aes(x = education, y = vocabulary)) +
  geom_jitter(alpha = 0.2)
p

# Plot 2: Add stat_sum
p +
  stat_sum() # sum statistic

# Plot 3: Set size range
p +
  stat_sum() + # sum statistic
  scale_size(range = c(1, 10)) # set size scale

```

## Stats outside Geoms

1. calculating statistics

```{r}
set.seed(123) 
xx <- rnorm(100)
mean(xx) + (sd(xx)* c(-1, 1))

# Hmisc
# install.packages("Hmisc")
library(Hmisc) 
smean.sdl(xx, mult = 1)

# ggplot2
mean_sdl(xx, mult = 1)
```

2. stat_summary() of mean and sd

```{r}
# uses geom_pointrange() by default
ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1))

ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1),
  geom = "errorbar", width = 0.1)

# NOT RECOMMENDED
ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  stat_summary(fun.y = mean, geom = "bar", fill = "skyblue") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1),
  geom = "errorbar", width = 0.1)

```

3. 95% Confidence Interval

```{r}
ERR <- qt(0.975, length(xx) - 1) * (sd(xx) / sqrt(length(xx)))
mean(xx) + (ERR* c(-1, 1))

# Hmisc
smean.cl.normal(xx) 

# ggplot2
mean_cl_normal(xx) 
            
```

4. stat_summary() of CI

```{r}
# use any function, as long as output has expected format
ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  stat_summary(fun.data = mean_cl_normal, width = 0.1)
```

5. Other stat_summary functions

* stat_summary(): Summarise y values at distinct x values

* stat_function(): Compute y values from a function of x values

* stat_qq(): Perform calculations for a quantile-quantile plot

6. Normal distribution

```{r}
# Normal distribution
library(MASS) 
mam.new <- data.frame(body = log10(mammals$body)) 
ggplot(mam.new, aes(x = body)) +
  geom_histogram(aes( y = ..density..)) +
  geom_rug() +
  stat_function(fun = dnorm, colour = "red", arg = list(mean = mean(mam.new$body), sd = sd(mam.new$body)))

# QQ plot
mam.new$slope <- diff(quantile(mam.new$body, c(0.25, 0.75))) / diff(qnorm(c(0.25, 0.75)))
mam.new$int <- quantile(mam.new$body, 0.25) - mam.new$slope * qnorm(0.25) 
ggplot(mam.new, aes(sample = body)) +
  stat_qq() +
  geom_abline(aes(slope = slope, intercept = int), col = "red")

```


### Practice

We'll set three position objects with convenient names. This allows us to use the exact positions on multiple layers. Create:

1. posn.d, using position_dodge() with a width of 0.1,

2. posn.jd, using position_jitterdodge() with a jitter.width of 0.1 and a dodge.width of 0.2

3. posn.j, using position_jitter() with a width of 0.2.

```{r}
# Display structure of mtcars
str(mtcars)

# Convert cyl and am to factors
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$am <- as.factor(mtcars$am)

# Define positions
posn.d <- position_dodge(width = 0.1)
posn.jd <- position_jitterdodge(jitter.width = 0.1, dodge.width = 0.2)
posn.j <- position_jitter(width = 0.2)

# Base layers
wt.cyl.am <- ggplot(mtcars, aes(x = cyl, y = wt,col = am, fill = am, group = am))
```



```{r}
# wt.cyl.am, posn.d, posn.jd and posn.j are available

# Plot 1: Jittered, dodged scatter plot with transparent points
wt.cyl.am +
  geom_point(position = posn.jd, alpha = 0.6)

# Plot 2: Mean and SD - the easy way
wt.cyl.am +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), position = posn.d)

# Plot 3: Mean and 95% CI - the easy way
wt.cyl.am +
  stat_summary(fun.data = mean_cl_normal, position = posn.d)

# Plot 4: Mean and SD - with T-tipped error bars - fill in ___
wt.cyl.am +
  stat_summary(geom = "point", fun.y = mean,
               position = posn.d) +
  stat_summary(geom = "errorbar", fun.data = mean_sdl,
               position = posn.d, fun.args = list(mult = 1), width = 0.1)
```



```{r}
# Play vector xx is available
mean_sdl(xx, mult = 1)

# Function to save range for use in ggplot
gg_range <- function(x) {
  # Change x below to return the instructed values
  data.frame(ymin = min(x), # Min
             ymax = max(x)) # Max
}

gg_range(xx)

# Function to Custom function
med_IQR <- function(x) {
  # Change x below to return the instructed values
  data.frame(y = median(x), # Median
             ymin = quantile(x)[2], # 1st quartile
             ymax = quantile(x)[4])  # 3rd quartile
}

med_IQR(xx)
```

When using stat_summary() recall that the fun.data argument requires a properly labelled 3-element long vector. The fun.y argument requires only a 1-element long vector.


1. The first stat_summary() layer should have geom set to "linerange". fun.data argument should be set to med_IQR, the function you used in the previous exercise.

2. The second stat_summary() layer also uses the "linerange" geom. This time fun.data should be gg_range, the other function you created. Also set alpha = 0.4.

3.For the last stat_summary() layer, use geom = "point". The points should have col "black" and shape "X".

```{r}
# The base ggplot command; you don't have to change this
wt.cyl.am <- ggplot(mtcars, aes(x = cyl,y = wt, col = am, fill = am, group = am))

# Add three stat_summary calls to wt.cyl.am
wt.cyl.am +
  stat_summary(geom = "linerange", fun.data = med_IQR,
               position = posn.d, size = 3) +
  stat_summary(geom = "linerange", fun.data = gg_range,
               position = posn.d, size = 3,
               alpha = 0.4) +
  stat_summary(geom = "point", fun.y = median,
               position = posn.d, size = 3,
               col = "black", shape = "X")
```

