---
title: 'Datacamp_Importing & Cleaning Data in R: Case Studies__School Attendance Data'
author: "dizhen"
date: "2019Äê4ÔÂ5ÈÕ"
output: html_document
---
# Importing the data
In this chapter, you'll work with attendance data from public schools in the US, organized by school level and state, during the 2007-2008 academic year. The data contain information on average daily attendance (ADA) as a percentage of total enrollment, school day length, and school year length.

The data were given to you in an Excel spreadsheet, which you can download or view a screenshot of. Either way, take a moment to look at the spreadsheet.

Do you see any symptoms of untidy data? At first glance, it looks like the first row is a description of the data, the second row is a variable itself that groups multiple columns together, and the fourth row gives numbers for the columns, which might look nice in a spreadsheet but isn't very useful for you, the analyst.

You'll take it one step at a time to import the data using the gdata package. The name of this spreadsheet is "attendance.xls" and is available in your working directory.

Instructions

1. Load the gdata package.
2. Read in the .xls file and store it as an object called att.

```{r}
# Load the gdata package
library(gdata)

# Import the spreadsheet: att
att <- read.xls("attendance.xls")
```
# Examining the data
For your reference, here is an image of the spreadsheet you are responsible for. Now that you've successfully imported the data into R, you can get a better idea of how the spreadsheet was read in.

Instructions

1. Print the column names of att.
2. Print the first 6 rows of att.
3. Print the last 6 rows of att.
4. Print the structure of att to get some information about your data.

```{r}
# Print the column names 
names(att)

# Print the first 6 rows
head(att,n = 6)

# Print the last 6 rows
tail(att, n = 6)

# Print the structure
str(att)
```

# Removing unnecessary rows
Again, for reference, here is an image of the first 22 rows of the original spreadsheet you were given. Looking at this image, you can see that rows 1, 4, 11, and 17 of the spreadsheet are useless. But is it safe to do something like the following?

att2 <- att[-c(1, 4, 11, 17), ]
No! From the last exercise, you might have realized that the read.xls() function actually imported the first row of the original data frame as the variable name for the first column. Did you notice that the first 6 rows of att aren't the same as the first six rows you saw in the original spreadsheet? What about the 11th and 17th rows?

When you're importing a messy spreadsheet into R, it's good practice to compare the original spreadsheet with what you've imported. It turns out that, by default, the read.xls() function skips empty rows such as the 11th and 17th.

After viewing your data frame, you realize you still need to get rid of the third row of att, as well as rows 56 through 59.

Instructions

1. Create a vector of row numbers to be removed. Call it remove.
2. Create att2 by removing these rows from att.

```{r}
# Create remove
remove <- c(3,56:59)

# Create att2
att2 <- att[-remove,]
```

# Removing useless columns
Once more, for reference, here is an image of the first 22 rows of the original spreadsheet. You can see here that the columns 3, 5, 7, 9, 11, 13, 15, and 17 (or columns C, E, G, I, K, M, O, Q in Excel) don't contain the values of average daily attendance (ADA). You'll get rid of them in this exercise.

```{r}
# Create remove
remove <- c( 3, 5, 7, 9, 11, 13, 15, 17)

# Create att3
att3 <- att2[,-remove]
```
# Splitting the data
In many cases, a single data frame stores multiple "tables" of information. You can often diagnose this problem by looking at the column names and noticing duplicate rows.

In this data frame, columns 1, 6, and 7 represent attendance data for US elementary schools, columns 1, 8, and 9 represent data for secondary schools, and columns 1 through 5 represent data for all schools in the US.

Each of these should be stored as its own separate data frame, so you'll split them up here.

att3 is pre-loaded in your workspace.

Instructions

1. Subset att3 to include only data for elementary schools (columns 1, 6, and 7). 2. Name the resulting data frame att_elem.
3. Subset att3 to include only data for secondary schools (columns 1, 8, and 9). 4. Name the resulting data frame att_sec.
5. Subset att3 to include data for all schools (columns 1 through 5). Name the resulting data frame att4.

```{r}
# Subset just elementary schools: att_elem
att_elem <- att3[,c(1,6,7)]

# Subset just secondary schools: att_sec
att_sec <- att3[,c(1,8,9)]

# Subset all schools: att4
att4 <- att3[,c(1:5)]
```
# Replacing the names
Since you went through so much trouble finding out which row stored the variable names, you should store that row as the actual column names of the data frame. We've modified the names a bit in order to be more stylistically sound; they're stored as cnames in the editor.

This will also allow you to remove the first two rows (currently storing variable names).

att4 is pre-loaded in your workspace.

Instructions

1. Run the first line of code as-is to define the cnames vector.
2. Assign the cnames vector to the column names of att4.
3. Remove the first two rows of att4 and name the result att5.
4. View the column names of att5.

```{r}
# Define cnames vector (don't change)
cnames <- c("state", "avg_attend_pct", "avg_hr_per_day", 
            "avg_day_per_yr", "avg_hr_per_yr")

# Assign column names of att4
colnames(att4) <- cnames

# Remove first two rows of att4: att5
att5<- att4[-c(1:2),]

# View the names of att5
names(att5)
```

# Cleaning up extra characters
One of the most irritating things about this dataset is that the state names are all stored as the same number of characters, with periods padding the ends of the shorter states. That may be helpful for reading the spreadsheet, but it makes your life harder, so you'll deal with it in this exercise.

One pitfall to avoid: . is a special character in the language of regular expressions (a.k.a. regex). In order to specify that you actually want to remove periods and not their regex equivalent (which is "all characters"), use \\.. This is called an "escape" sequence.

stringr and att5 are pre-loaded in your workspace.

Instructions

1. Use the function str_replace_all() to replace all periods in the state column of att5 with "". Remember to use \\. to represent a period. Assign the result back to att5$state.
2. Remove white space around the state names, assigning the result back to att5$state once more. There's another stringr function called str_trim() for this purpose.
3. View the head of att5.

```{r}
# Remove all periods in state column
att5$state <- str_replace_all(att5$state,"\\.","")

# Remove white space around state names
att5$state <- str_trim(att5$state)

# View the head of att5
head(att5)
```

# Some final type conversions
Finally, you'll convert the values in certain variables to numerics (instead of factors). It's worth noting that in previous chapters, your numerical data has often come in as character strings. This is just a difference between read.xls() and the other import functions you've used.

The dplyr package offers an efficient method for applying a function to many columns at once. If you'd like to learn more, check out DataCamp's course on Data Manipulation in R with dplyr! Since you might not have taken the dplyr course yet, we're just showing you the code here. You'll do the same thing, but using subsetting and sapply() instead.

Instructions

1. Review the dplyr code, just to see how it works.
2. Define a vector called cols that contains indices of columns containing numerical data. Take a look at att5 if you need to!
3. Using a call to sapply(), single bracket subsetting, and the vector cols, change the type of these columns to numeric with as.numeric(). Assign the result of the function call back to the subset of att5.

```{r}
# Change columns to numeric using dplyr (don't change)
library(dplyr)
example <- mutate_at(att5, vars(-state), funs(as.numeric))

# Define vector containing numerical columns: cols
cols <- -1

# Use sapply to coerce cols to numeric
att5[, cols] <- sapply(att5[, cols], as.numeric)
```

