---
title: "Datacamp_Data Manipulation in R with dplyr__Group_by and working with databases"
author: "dizhen"
date: "2019Äê4ÔÂ6ÈÕ"
output: html_document
---
# Unite and conquer using group_by
As Garrett explained, group_by() lets you define groups within your data set. Its influence becomes clear when calling summarize() on a grouped dataset: summarizing statistics are calculated for the different groups separately.

In this exercise, you are going to create an ordered per-carrier summary of hflights by combining group_by(), summarize() and arrange().

Instructions

1. Use group_by() to group hflights by UniqueCarrier.

2. summarize() the grouped tbl with two summary variables:

p_canc, the percentage of cancelled flights.
avg_delay, the average arrival delay of flights whose delay does not equal NA.

3. Finally, order the carriers in the summary from low to high by their average arrival delay. Use percentage of flights cancelled to break any ties.

```{r}
# hflights is in the workspace as a tbl, with translated carrier names

# Make an ordered per-carrier summary of hflights
hflights %>%
  group_by(UniqueCarrier) %>%
  summarize(
    p_canc = mean(Cancelled==1)*100,
    avg_delay = mean(ArrDelay,na.rm=TRUE)
  ) %>%
  arrange(avg_delay, p_canc)
```
# Combine group_by with mutate
You can also combine group_by() with mutate(). When you mutate grouped data, mutate() will calculate the new variables independently for each group. This is particularly useful when mutate() uses the rank() function, that calculates within-group rankings. rank() takes a group of values and calculates the rank of each value within the group, e.g.

rank(c(21, 22, 24, 23))

has output

[1] 1 2 4 3

As with arrange(), rank() ranks values from the smallest to the largest.

Instructions

1. filter() the hflights tbl to only keep observations for which ArrDelay is not NA and positive.

2. Use group_by() on the result to group by UniqueCarrier.

3. Next, use summarize() to calculate the average ArrDelay per carrier. Call this summary variable avg.

4. Feed the result into a mutate() call: create a new variable, rank, calculated as rank(avg).

5. Finally, arrange by this new rank variable

```{r}
# dplyr is loaded, hflights is loaded with translated carrier names

# Ordered overview of average arrival delays per carrier
hflights %>% filter(!is.na(ArrDelay), ArrDelay > 0) %>% group_by(UniqueCarrier) %>% summarize(avg = mean(ArrDelay)) %>% mutate(rank = rank(avg)) %>% arrange(rank)
```

# Advanced group_by exercises
By now you've learned the fundamentals of dplyr: the five data manipulation verbs and the additional group_by() function to discover interesting group-wise statistics. The next challenges are an all-encompassing review of the concepts you have learned about. We already provided you with a template of the piped call that can solve the exercises. Up to you to finish all dplyr calls! For simplicity, you can include cancelled flights in your answers, so you shouldn't filter based on the Cancelled column.

Instructions

1. How many airplanes flew to only one destination? The tbl you print out should have a single column, named nplanes and a single row.

2. Find the most visited destination for each carrier. The tbl you print out should contain four columns:

UniqueCarrier and Dest,

n, how often a carrier visited a particular destination,

rank, how each destination ranks per carrier. rank should be 1 for every row, as you want to find the most visited destination for each carrier.


```{r}
# dplyr and hflights (with translated carrier names) are pre-loaded

# How many airplanes only flew to one destination?
hflights %>%
  group_by(TailNum) %>%
  summarize(ndest = n_distinct(Dest)) %>%
  filter(ndest == 1) %>%
  summarize(nplanes = n())

# Find the most visited destination for each carrier
hflights %>%
  group_by(UniqueCarrier, Dest) %>%
  summarize(n = n()) %>%
  mutate(rank = rank(desc(n))) %>%
  filter(rank == 1)
```

# dplyr deals with different types
hflights2 is a copy of hflights that is saved as a data table. hflights2 was made available in the background using the following code:

library(data.table)
hflights2 <- as.data.table(hflights)
hflights2 contains all of the same information as hflights, but the information is stored in a different data structure. You can see this structure by typing hflights2 at the command line.

Even though hflights2 is a different data structure, you can use the same dplyr functions to manipulate hflights2 as you used to manipulate hflights.

Instructions

Use summarize() to calculate n_carrier, the total number of unique carriers in hflights2. Whether or not you use the pipe is up to you!

```{r}
# hflights2 is pre-loaded as a data.table

# Use summarize to calculate n_carrier
hflights2 %>% summarize(n_carrier = n_distinct(UniqueCarrier))
```

# dplyr and mySQL databases

DataCamp hosts a mySQL database with data about flights that departed from New York City in 2013. The data is similar to the data in hflights, but it does not contain information about cancellations or diversions. With the tbl() function, we already created a reference to a table in this information.

Although nycflights is a reference to data that lives outside of R, you can use the dplyr commands on them as usual. Behind the scenes, dplyr will convert the commands to the database's native language (in this case, SQL), and return the results. This allows you to pull data that is too large to fit in R: only the fraction of the data that you need will actually be downloaded into R, which will usually fit into R without memory issues.

Instructions

1. Try to understand the code that creates nycflights, a reference to a MySQL table.

2. Use glimpse() to check out nycflights. Although nycflights is a reference to a tbl in a remote database, there is no difference in syntax. Look carefully: the variable names in nycflights differ from the ones in hflights!

3. Group nycflights data by carrier, then summarize() with two variables: n_flights, the number of flights flown by each carrier and avg_delay, the average arrival delay of flights flown by each carrier. Finally, arrange the carriers by average delay from low to high.

```{r}
# Set up a connection to the mysql database
my_db <- src_mysql(dbname = "dplyr", 
                   host = "courses.csrrinzqubik.us-east-1.rds.amazonaws.com", 
                   port = 3306, 
                   user = "student",
                   password = "datacamp")

# Reference a table within that source: nycflights
nycflights <- tbl(my_db, "dplyr")

# glimpse at nycflights
glimpse(nycflights)

# Ordered, grouped summary of nycflights
nycflights %>% group_by(carrier) %>% summarize(n_flights = n(),avg_delay = mean(arr_delay)) %>% arrange(avg_delay)
```

