---
title: "Datacamp_Data Manipulation in R with dplyr__Select and mutate"
author: "dizhen"
date: "2019Äê4ÔÂ5ÈÕ"
output: html_document
---
# The five verbs and their meaning
The dplyr package contains five key data manipulation functions, also called verbs:

select(), which returns a subset of the columns,
filter(), that is able to return a subset of the rows,
arrange(), that reorders the rows according to single or multiple variables,
mutate(), used to add columns from existing data,
summarize(), which reduces each group to a single row by calculating aggregate measures.
If you want to find out more about these functions, consult the documentation by clicking on the functions above. What order of operations should we use to find the average value of the ArrDelay (arrival delay) variable for all American Airline flights in the hflights tbl?

Feel free to play around in the console; hflights is preloaded. From now on, the UniqueCarrier column and CancellationCode column contain the recoded versions, similar to the cleaning up you did in the previous chapter.

Answer:

First filter(), then summarize().

# Choosing is not losing! The select verb
To answer the simple question whether flight delays tend to shrink or grow during a flight, we can safely discard a lot of the variables of each flight. To select only the ones that matter, we can use select().

As an example, take the following call, that selects the variables var1 and var2 from the data frame df.

select(df, var1, var2)
You can also use : to select a range of variables and - to exclude some variables, similar to indexing a data.frame with square brackets. You can use both variable's names as well as integer indexes. This call selects the four first variables except for the second one of a data frame df:

select(df, 1:4, -2)
select() does not change the data frame it is called on; you have to explicitly assign the result of select() to a variable to store the result.

Instructions

1. Use select() to print out a tbl that contains only the columns ActualElapsedTime, AirTime, ArrDelay and DepDelay of hflights.
2. Print out a tbl with the columns Origin up to and including Cancelled of hflights.
3. Find the most concise way to select: columns Year up to and including DayOfWeek, columns ArrDelay up to and including Diverted. You can examine the order of the variables in hflights with names(hflights) in the console.

```{r}
# hflights is pre-loaded as a tbl, together with the necessary libraries.

# Print out a tbl with the four columns of hflights related to delay
select(hflights, ActualElapsedTime, AirTime, ArrDelay, DepDelay)

# Print out the columns Origin up to Cancelled of hflights
select(hflights,Origin:Cancelled)

# Answer to last question: be concise!
select(hflights,1:4,12:21)
```

# Helper functions for variable selection
dplyr comes with a set of helper functions that can help you select groups of variables inside a select() call:

starts_with("X"): every name that starts with "X",
ends_with("X"): every name that ends with "X",
contains("X"): every name that contains "X",
matches("X"): every name that matches "X", where "X" can be a regular expression,
num_range("x", 1:5): the variables named x01, x02, x03, x04 and x05,
one_of(x): every name that appears in x, which should be a character vector.
Pay attention here: When you refer to columns directly inside select(), you don't use quotes. If you use the helper functions, you do use quotes.

Instructions

1. Use select() and a helper function to print out a tbl that contains just ArrDelay and DepDelay of hflights.
2. Use a combination of helper functions and variable names to print out only the UniqueCarrier, FlightNum, TailNum, Cancelled, and CancellationCode columns of hflights.
3. Find the most concise way to return the following columns with select and its helper functions: DepTime, ArrTime, ActualElapsedTime, AirTime, ArrDelay, DepDelay. Use only helper functions!

```{r}
# As usual, hflights is pre-loaded as a tbl, together with the necessary libraries.

# Print out a tbl containing just ArrDelay and DepDelay
select(hflights,ends_with("Delay"))

# Print out a tbl as described in the second instruction, using both helper functions and variable names
select(hflights,UniqueCarrier,ends_with("Num"),starts_with("Cancell"))

# Print out a tbl as described in the third instruction, using only helper functions.
select(hflights,ends_with("Time"),ends_with("Delay"))
```
# Comparison to base R
To see the added value of the dplyr package, it is useful to compare its syntax with base R. Up to now, you have only considered functionality that is also available without the use of dplyr. The elegance and ease-of-use of dplyr is a great plus though.

Instructions

1. Finish the select() calls to match the results of the base R commands. Try to make your calls as concise as possible.

```{r}
# both hflights and dplyr are available

# Finish select call so that ex1d matches ex1r
ex1r <- hflights[c("TaxiIn", "TaxiOut", "Distance")]
ex1d <- select(hflights,starts_with("Taxi"),Distance)

# Finish select call so that ex2d matches ex2r
ex2r <- hflights[c("Year", "Month", "DayOfWeek", "DepTime", "ArrTime")]
ex2d <- select(hflights,1:6,-3)

# Finish select call so that ex3d matches ex3r
ex3r <- hflights[c("TailNum", "TaxiIn", "TaxiOut")]
ex3d <- select(hflights,starts_with("Ta"))

```
# Mutating is creating
mutate() is the second of five data manipulation functions you will get familiar with in this course. mutate() creates new columns which are added to a copy of the dataset.

Take this example that adds a new column, z, which is the element-wise sum of the columns x and y, to the data frame df:

mutate(df, z = x + y)

Instructions

1. Create a new data frame, g1, which is the data frame hflights with an additional column: ActualGroundTime, the difference between ActualElapsedTime and AirTime.
2. Extend g1 further, by adding an additional column GroundTime. This column is the sum of the TaxiIn and TaxiOut columns. Store the resulting data frame in g2. Check in the console that the GroundTime and ActualGroundTime columns are equal.
3. Add a new variable to g2 named AverageSpeed that denotes the average speed that each plane flew in miles per hour. Save the resulting dataset as g3. Use the following formula: 60 * Distance / AirTime.
4. Print out g3.

```{r}
# hflights and dplyr are loaded and ready to serve you.

# Add the new variable ActualGroundTime to a copy of hflights and save the result as g1.
g1 <- mutate(hflights,ActualGroundTime =ActualElapsedTime-AirTime)

# Add the new variable GroundTime to g1. Save the result as g2.
g2 <- mutate(g1, GroundTime = TaxiIn + TaxiOut)

# Add the new variable AverageSpeed to g2. Save the result as g3.
g3 <- mutate(g2, AverageSpeed = 60 * Distance / AirTime)

# Print out g3
g3
```
# Add multiple variables using mutate
So far you've added variables to hflights one at a time, but you can also use mutate() to add multiple variables at once. To create more than one variable, place a comma between each variable that you define inside mutate().

mutate() even allows you to use a new variable while creating a next variable in the same call. In this example, the new variable x is directly reused to create the new variable y:

mutate(my_df, x = a + b, y = x + c)

Instructions

1. Adapt the code that builds m1: add a variable loss_ratio, which is the ratio of loss to DepDelay.
2. Create a tbl m2 from hflights by adding three variables:
3. TotalTaxi, which is the sum of TaxiIn and TaxiOut;
4. ActualGroundTime, which is the difference of ActualElapsedTime and AirTime;
5. Diff, the difference between the two newly created variables. This column should be zero for all observations!

```{r}
# hflights and dplyr are ready, are you?

# Add a second variable loss_ratio to the dataset: m1
m1 <- mutate(hflights, loss = ArrDelay - DepDelay,loss_ratio = loss/DepDelay)

# Add the three variables as described in the third instruction: m2
m2 <- mutate(hflights, TotalTaxi = TaxiIn + TaxiOut, ActualGroundTime = ActualElapsedTime - AirTime, Diff = TotalTaxi - ActualGroundTime)

```
# Recap on mutate and select
As of now, you mastered two of the five data manipulation functions that are at the core of dplyr: select() and mutate().

Which statement concerning the following four expressions is correct?

(A) hflights <- select(hflights, -(Year:Month), -(DepTime:Diverted))

(B) select(hflights, starts_with("D"))

(C) select(hflights, -(Year:Month), -(DepTime:Diverted))

(D) hflights <- select(hflights, starts_with("Day"))

Feel free to experiment in the console, hflights is loaded as a tbl.


