---
title: "Datacamp_Grouping_and_Summarizing"
author: "dizhen"
date: "2019/4/4"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
gapminder <- read.table(file = 'data/gapminder.tsv', sep = '\t', header = TRUE)
```

```{r}
# Extracting data
gapminder %>% filter(country == "United States", year == 2007) %>% head()
```

## The summarize verb

summarize() turns many rows into one

```{r}
gapminder %>% summarize(meanLifeExp = mean(lifeExp))
```
```{r}
gapminder %>% filter(year == 2007) %>% summarize(meanLifeExp = mean(lifeExp))
```

```{r}
gapminder %>%
  filter(year == 2007) %>%
  summarize(meanLifeExp = mean(lifeExp), totalPop = sum(as.numeric(pop)))
```
Functions you can use for sumarizing:

* mean, sum, median, min, max


```{r}

# Summarize to find the median life expectancy
gapminder %>%
  summarize(medianLifeExp = median(lifeExp))
```



```{r}
# Filter for 1957 then summarize the median life expectancy
gapminder %>%
  filter(year == 1957) %>%
  summarize(medianLifeExp = median(lifeExp))
```

```{r}

# Filter for 1957 then summarize the median life expectancy and the maximum GDP per capita
gapminder %>% filter(year == 1957) %>% summarize(medianLifeExp = median(lifeExp),maxGdpPercap = max(gdpPercap))
```

## The group_by verb

group_by() before summarize() turns groups into one row each.


```{r}
gapminder %>% 
  filter(year == 2007) %>%
  summarize(meanLifeExp = mean(lifeExp), totalPop = sum(as.numeric(pop)))
```



```{r}
# Summarizing by year
gapminder %>% 
  group_by(year) %>%
  summarize(meanLifeExp = mean(lifeExp), totalPop = sum(pop)) %>%head()
```
```{r}
# Summarizing by continent
gapminder %>% 
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarize(meanLifeExp = mean(lifeExp), totalPop = sum(pop))
```

```{r}
gapminder %>%
  group_by(year, continent) %>%
  summarize(totalPop = sum(pop), meanLifeExp = mean(lifeExp)) %>%head()
```

### Practice

```{r}

# Find median life expectancy and maximum GDP per capita in each year
gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp),
            maxGdpPercap = max(gdpPercap)) %>%head()
```

Summarizing by continent


```{r}

# Find median life expectancy and maximum GDP per capita in each continent in 1957
gapminder %>%
  filter(year == 1957) %>%
  group_by(continent) %>%
  summarize(medianLifeExp = median(lifeExp),
            maxGdpPercap = max(gdpPercap))
```

Summarizing by continent and year


```{r}

# Find median life expectancy and maximum GDP per capita in each continent/year combination
gapminder %>%
  group_by(continent, year) %>%
  summarize(medianLifeExp = median(lifeExp),
            maxGdpPercap = max(gdpPercap)) %>%head()
```



## Visualizing summarized data

```{r}
# Summarizing by year
by_year <- gapminder %>%
  group_by(year) %>%
  summarize(totalPop = sum(as.numeric(pop)),meanLifeExp = mean(lifeExp))

ggplot(by_year, aes(x = year, y = totalPop)) +
  geom_point()
```

```{r}
# Starting y-axis at zero
ggplot(by_year, aes(x = year, y = totalPop)) + 
  geom_point() +
  expand_limits(y = 0)
```
```{r}
# Summarizing by year and continent
by_year_continent <- gapminder %>% 
  group_by(year, continent) %>%
  summarize(totalPop = sum(as.numeric(pop)), meanLifeExp = mean(lifeExp))

ggplot(by_year_continent, aes(x = year, y = totalPop, color = continent)) +
  geom_point() +
  expand_limits(y = 0)
```


### Practice

Visualizing median life expectancy over time

```{r}

by_year <- gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp),
            maxGdpPercap = max(gdpPercap))

# Create a scatter plot showing the change in medianLifeExp over time
ggplot(by_year,aes(x = year, y = medianLifeExp)) +
  geom_point() +
  expand_limits(y = 0)
```

Visualizing median GDP per capita per continent over time


```{r}

# Summarize medianGdpPercap within each continent within each year: by_year_continent
by_year_continent <- gapminder %>% group_by(continent,year) %>% summarize(medianGdpPercap = median(gdpPercap))

# Plot the change in medianGdpPercap in each continent over time
ggplot(by_year_continent,aes(x = year , y = medianGdpPercap, color = continent))+
  geom_point()+
  expand_limits(y = 0)
```

Comparing median life expectancy and median GDP per continent in 2007


```{r}

# Summarize the median GDP and median life expectancy per continent in 2007
by_continent_2007 <- gapminder %>% filter(year == 2007) %>% group_by(continent) %>% summarize(medianLifeExp = median(lifeExp), medianGdpPercap = median(gdpPercap))

# Use a scatter plot to compare the median GDP and median life expectancy
ggplot(by_continent_2007, aes(x = medianGdpPercap, y = medianLifeExp, color = continent)) + geom_point() + expand_limits(y = 0)
```

