---
title: "Datacamp_Joining Data in R with dplyr__Case study"
author: "dizhen"
date: "2019Äê4ÔÂ7ÈÕ"
output: html_document
---
# Universal keys?
As Garrett mentioned, the Lahman database contains 26 datasets. Before starting to analyze the data, you'll examine how the datasets are related to each other.

In particular, do any variable names span all of the datasets?

Instructions

1. Examine lahmanNames by printing it to the console. It is a list of 26 data frames, each of which contains the variable names of one dataset in the Lahman package.

2. Use purrr's reduce() and a single set operation to return a data frame that contains every variable that appears in all 26 datasets (if any does).

```{r}
# Examine lahmanNames
lahmanNames

# Find variables in common
reduce(lahmanNames,intersect)
```

# Common keys
No variables span all 26 datasets, but several variables span more than one dataset. These variables provide relationships between the various tables in Lahman. Can you find them?

Instructions

1. Bind the data frames in lahmanNames into a single table that includes a new column called dataframe, which contains the names of the data frames.

2. Group the result by var.

3. Tally the number of appearances of each var.

4. Filter the data to only variables that appear more than once.

5. Arrange the results so that the most frequently used variables appear before less frequently used variables.

```{r}
lahmanNames %>%  
  # Bind the data frames in lahmanNames
  bind_rows(.id = "dataframe") %>%
  # Group the result by var
  group_by(var) %>%
  # Tally the number of appearances
  tally() %>%
  # Filter the data
  filter(n > 1) %>% 
  # Arrange the results
  arrange(desc(n))
```

# playerID
In the last exercise, you saw that playerID is one of the most widely used keys in the Lahman database. Which datasets use playerID?

Instructions

1. Bind the data frames in lahmanNames into a single table that includes the column dataframe, which contains the names of the data frames.

2. Filter the results to just rows that contain playerID.

3. Extract the dataframe variable.

```{r}
lahmanNames %>% 
  # Bind the data frames
  bind_rows(.id = "dataframe") %>%
  # Filter the results
  filter(var == "playerID") %>% 
  # Extract the dataframe variable
  `$`(dataframe)
```

# Who are the players?
Before we look at salaries, let's begin by ensuring that we have salary information for each player in the database, or at least no systematic holes in our coverage.

To do this we will need a list of every player to compare against salaries. The Master dataset contains all of the players in the database, but it may contain multiple rows for each player.

How should you proceed?

In the video, you saw how to use distinct() to find unique rows for all columns in a table. You can also find unique rows for specific columns using the following syntax:

```{r}
tbl %>%
  # Find unique rows of columns a,b, and c
  distinct(a, b, c)
```

Instructions

1. Use Master to create players, a data frame that contains one row and three columns for each distinct player. The columns should be playerID, nameFirst, and nameLast, in that order.

```{r}
players <- Master %>% 
  # Return one row for each distinct player
  distinct(playerID,nameFirst, nameLast)
```

# Missing salaries
Now that we know who the players are, let's check our salary coverage. How many players are completely missing salary information?

Instructions

1. Use a join to find all of the players who do not appear in Salaries.

2. Then use the dplyr function count() to count how many players do not appear in Salaries.

```{r}
players %>% 
  # Find all players who do not appear in Salaries
  anti_join(Salaries) %>%
  # Count them
  count()
```

# Unpaid games?
Now that we know there is a huge hole in the salary data, let's see if we can explain it. Is it possible that these players somehow did not play (and hence did not earn a salary)?

We can check with the Appearances data frame. Appearances contains information about every game played in major league baseball. That is, if a player played a game, it would show up as a row in Appearances.

Instructions

Add a join to the pipe to determine how many (if any) unsalaried players played at least one game.

```{r}
players %>% 
  anti_join(Salaries, by = "playerID") %>% 
  # How many unsalaried players appear in Appearances?
  semi_join(Appearances, by = "playerID") %>% 
  count()

```

# How many games?
Perhaps the unsalaried players only played one or two games, and hence did not earn a full salary. Can you determine how many games each of these unsalaried players played?

Instructions

1. Recalculate the list of players who do not appear in Salaries with the anti_join() function.

2. Then left join them to Appearances.

3. Group the data by playerID and calculate the total_games that each player played. This will be the sum() of G_all, ignoring NAs .

4. Arrange the summarized data in descending order by total_games.

```{r}
players %>% 
  # Find all players who do not appear in Salaries
  anti_join(Salaries, by = "playerID") %>% 
  # Join them to Appearances
  left_join(Appearances, by = "playerID") %>% 
  # Calculate total_games for each player
  group_by(playerID) %>%
  summarize(total_games = sum(G_all, na.rm = TRUE)) %>%
  # Arrange in descending order by total_games
  arrange(desc(total_games))
```

# How many at-bats?
Is it possible that the unsalaried players did not actually play in the games that they appeared in? One way to check would be to determine if the players had an at-bat (i.e. batted) in the games that they appeared in.

Instructions

Repeat your process in the previous exercise with Batting instead of Appearances:

1. Find the unsalaried players.

2. Left join Batting to the unsalaried players.

3. Group by player (e.g. playerID).

4. Calculate the sum of at-bats for each player as total_at_bat; this will be the sum of the AB variable in Batting, ignoring all NAs.

5. Arrange in descending order of at-bats.

```{r}
players %>%
  # Find unsalaried players
  anti_join(Salaries, by = "playerID") %>% 
  # Join Batting to the unsalaried players
  left_join(Batting, by = "playerID") %>% 
  # Group by player
  group_by(playerID) %>% 
  # Sum at-bats for each player
  summarize(total_at_bat = sum(AB, na.rm = TRUE)) %>% 
  # Arrange in descending order
  arrange(desc(total_at_bat))
```

# Hall of fame nominations
As Garrett mentioned, the Hall of Fame is a collection of distinguished baseball players selected by a committee of baseball experts. As with any hall of fame, more players are nominated for membership than are actually admitted.

Let's see how many players have been nominated for the Hall of Fame.

Instructions

1. Create nominated, a data frame that contains the distinct players (by playerID) that appear in the HallOfFame dataset.

2. Count the number of players in nominated with count().

3. Add nameFirst and nameLast values from the Master data frame to nominated. Store the result to a data frame called nominated_full.

```{r}
# Find the distinct players that appear in HallOfFame
nominated <- HallOfFame %>% 
  distinct(playerID)

nominated %>% 
  # Count the number of players in nominated
  count()

nominated_full <- nominated %>% 
  # Join to Master
  left_join(Master[,c("playerID","nameFirst","nameLast")],by = "playerID")%>% 
  # Return playerID, nameFirst, nameLast
  select(playerID,nameFirst,nameLast)
```

# Hall of fame inductions
In the previous exercise, we saw that 1,239 players were nominated for the hall of fame. Let's now see how many players were admitted to the hall of fame to examine how selective the voting process is.

Instructions

1. Repeat the process you saw in the last exercise to create inducted, a data frame that contains the distinct players (by playerID) that appear in HallOfFame and have the value inducted == "Y".

2£¬Count the number of players in inducted with count().

3. Add nameFirst and nameLast values from the Master data frame to inducted. Store the result to a data frame called inducted_full.

```{r}
# Find distinct players in HallOfFame with inducted == "Y"
inducted <- HallOfFame %>% 
  filter(inducted == 'Y') %>% 
  distinct(playerID)

inducted %>% 
  # Count the number of players in inducted
  count()

inducted_full <- inducted %>% 
  # Join to Master
  left_join(Master[,c("playerID","nameFirst","nameLast")], by = "playerID") %>% 
  # Return playerID, nameFirst, nameLast
  select(playerID, nameFirst, nameLast)
```

# Awards
Now that we know who was inducted and who was nominated, let's examine what separates the nominees who were inducted from the nominees who were not.

Let's start with a simple question: Did nominees who were inducted earn more awards than nominees who were not inducted?

We can use AwardsPlayers to answer the question. It lists the playerID's of players who won baseball awards, and it contains one row for each award awarded in major league baseball.

Instructions

1. tally() the number of awards in AwardsPlayers for each playerID and save the results as nAwards. Note that the resulting column containing the number of awards is called n by default.

2. Filter nAwards to include the players in inducted and calculate the mean number of awards per player (set na.rm = TRUE). Call the mean number of awards avg_n.

3£¬Filter nAwards to include the players in nominated who are not present in inducted and calculate the mean number of awards per player (set na.rm = TRUE). Again, call the mean number of awards avg_n.

```{r}
# Tally the number of awards in AwardsPlayers by playerID
nAwards <- AwardsPlayers %>% 
  group_by(playerID) %>% 
  tally()

nAwards %>% 
  # Filter to just the players in inducted 
  semi_join(inducted,by = "playerID") %>% 
  # Calculate the mean number of awards per player
  summarize(avg_n = mean(n, na.rm = TRUE))

nAwards %>% 
  # Filter to just the players in nominated 
  semi_join(nominated,by = "playerID") %>% 
  # Filter to players NOT in inducted 
  anti_join(inducted,by = "playerID") %>% 
  # Calculate the mean number of awards per player
  summarize(avg_n = mean(n, na.rm = TRUE))
```

# Salary
Salary may provide another way to differentiate inductees from non-inductees. Does the maximum salary earned by inductees tend to be greater than the maximum salary earned by nominees who were not inducted?

Instructions

1. Create notInducted, a dataset that contains the rows of nominated that are not in inducted using a set operation.

2. Filter Salaries against notInducted to return only the salaries of nominees who were not inducted.
Then calculate the max salary for each player in the results. Call this variable max_salary.
Then in a separate summarize() call, calculate the average of the max salaries. Call this variable avg_salary.

3. Repeat the process for the players in inducted.

```{r}
# Find the players who are in nominated, but not inducted
notInducted <- nominated %>% 
  anti_join(inducted, by = "playerID") # setdiff() is more correct

Salaries %>% 
  # Find the players who are in notInducted
  semi_join(notInducted, by = "playerID") %>% 
  # Calculate the max salary by player
  group_by(playerID) %>% 
  summarize(max_salary = max(salary)) %>% 
  # Calculate the average of the max salaries
  summarize(avg_salary = mean(max_salary))

# Repeat for players who were inducted
Salaries %>% 
  semi_join(inducted, by = "playerID") %>% 
  group_by(playerID) %>% 
  summarize(max_salary = max(salary)) %>% 
  summarize(avg_salary = mean(max_salary))
```

# Retirement
One of the rules of the Hall of Fame is that players cannot be nominated until five years after they retire. Is this reflected in our data?

Instructions

Use Appearances to determine the year of the last appearance of each player nominated to the Hall of Fame. Compare it to the year(s) they were nominated to determine whether any players were nominated before they completely retired. You'll want to:

1. Filter Appearances against nominated to get the information in Appearances for all nominated players.

2. Use dplyr methods to find the last year played by player. Call this last_year within your summarize() call.

3. Left join the results to the full HallOfFame dataset.

4. Use dplyr methods to spot unusual observations.

```{r}
Appearances %>% 
  # Filter Appearances against nominated
  semi_join(nominated, by = "playerID") %>% 
  # Find last year played by player
  group_by(playerID) %>% 
  summarize(last_year = max(yearID, na.rm = TRUE)) %>% 
  # Join to full HallOfFame
  left_join(HallOfFame, by = 'playerID') %>% 
  # Filter for unusual observations
  filter(last_year >= yearID)
```

