---
title: "Datacamp_Introduction to Data__case study"
author: "dizhen"
date: "2019Äê4ÔÂ14ÈÕ"
output: html_document
---
# Identify variable types
It's always useful to start your exploration of a dataset by identifying variable types. The results from this exercise will help you design appropriate visualizations and calculate useful summary statistics later in your analysis.

Instructions

Explore the evals data once again with the following goals in mind: (1) Identify each variable as numerical or categorical. (2) If numerical, determine if it is discrete or continuous. (3) If categorical, determine if it is ordinal or not.
We've created a vector of variable names in the editor called cat_vars. To test your understanding of the data, remove the names of any variables that are not categorical.

```{r}
# Inspect variable types
str(evals) 
glimpse(evals)

# Remove non-factor variables from the vector below
cat_vars <- c("rank", "ethnicity", "gender", "language", "cls_level", "cls_profs", "cls_credits", "pic_outfit", "pic_color")
```

# Recode a variable
The cls_students variable in evals tells you the number of students in the class. Suppose instead of the exact number of students, you're interested in whether the class is

"small" (18 students or fewer),
"midsize" (19 - 59 students), or
"large" (60 students or more).

Instructions

Recode the cls_students variable into a new variable, cls_type, using case_when(). This new variable should have three levels as described above. Save the resulting data frame (with the new variable) as evals.
What type of variable is cls_type?

```{r}
x = case_when(
    [condition 1] ~ [value of x when condition 1 is true],
    [condition 2] ~ [value of x when condition 2 is true],
    [condition 3] ~ [value of x when condition 3 is true]
)
```

```{r}
# Recode cls_students as cls_type
evals_fortified <- evals %>%
  mutate(
    cls_type = case_when(
      cls_students <= 18                      ~ "small",
      cls_students >= 19 & cls_students <= 59 ~ "midsize",
      cls_students >= 60                      ~ "large"
    )
  )
```
# Create a scatterplot
The bty_avg variable shows the average beauty rating of the professor by the six students who were asked to rate the attractiveness of these faculty. The score variable shows the average professor evaluation score, with 1 being very unsatisfactory and 5 being excellent.

Instruction

Use ggplot() to create a scatterplot displaying the relationship between these two variables.
How would you describe the relationship apparent in this visualization?

```{r}
# Scatterplot of score vs. bty_avg
ggplot(evals, aes(x = bty_avg, y = score)) +
  geom_point()
```

# Create a scatterplot, with an added layer
Suppose you are interested in evaluating how the relationship between a professor's attractiveness and their evaluation score varies across different class types (small, midsize, and large).

Instructions

Recreate your visualization from the previous exercise, but this time coloring the points by class type.
How would you describe the relationship apparent in this visualization?

```{r}
# Scatterplot of score vs. bty_avg colored by cls_type
ggplot(evals, aes(x = bty_avg, y = score, color = cls_type))+
  geom_point()
```

